---
title: "SAMS Student Data Diagnostics"
output: html_document
date: "2024-08-29"
---
This document provides a diagnostics of SAMS data - API access - CSM. 



# Import the .parquest datafile
```{r}
install.packages("arrow")
library(arrow)

# Replace 'your_file.parquet' with the path to your Parquet file
df <- read_parquet("C:/Users/Admin/Box/Amitav/Skills/student_data.parquet")

table(df$module)

```



# Check for Duplicates in id variables
```{r}
# Check for the non-unique Barcode × Year. Generates a column which is a count of duplicates
df_duplicates_barcode <- df %>%
  group_by(Barcode,ReportedInstitute) %>%
  summarize(num_duplicates = n())

duplicate_barcodes <- df_duplicates_barcode %>%
 filter(num_duplicates > 1) %>%   # Filter rows where num_duplicates are greater than 1
 select(Barcode)           # Select only those Barcodes


barcode_dupli_list = df %>% filter(Barcode %in% duplicate_barcodes$Barcode)

write.xlsx(barcode_dupli_list, "barcode_duplicates_df.xlsx")

 
#filtered_sams = sams_student_data %>% filter(Barcode %in% duplicate_barcodes$Barcode)
#view(filtered_sams)
#remove(sams_student_data)
#print(df_duplicates_barcode$Barcode>1)

#duplicate_barcodes <- df_duplicates_barcode %>%
 # filter(num_duplicates > 1) %>%   # Filter rows where col1 is greater than 1
  #select(Barcode)           # Select only col1

```

# Check if all the institutes in 2024 (ITI, Diploma and PDIS) are present in the dataset
This should be done after the admission cycle has ended to check if the "college participation" figures for 2024 on SAMS website matches with the SAMS dataset. 
```{r}
## For ITIs
missing_itis = setdiff(gsub(" ", "", iti_list_2024_25$Institute_Name), gsub(" ", "", sams_student_data$ReportedInstitute))

print("Number of ITIs missing from SAMS")
print(missing_itis)

remove(d_list,e_list)
## For Diploma

## For PDIS

```



# Year-wise count
```{r}
# Calculate frequency of Gender and Year for all institutes
freq_df_gender <- sams_student_data %>%
  group_by(Year, Gender) %>%
  summarise(Frequency = n()) %>%
  mutate(Percentage = Frequency / sum(Frequency) * 100) %>%
  ungroup()

# Create the stacked bar chart
ggplot(freq_df_gender, aes(x = as.factor(Year), y = Frequency, fill = Gender)) +
  geom_bar(stat = "identity", color = "black") +  # Bar color and border
  labs(
    title = "Count SAMS Reported by Year and Gender",
    x = "Year",
    y = "Count",
    fill = "Gender"
  ) +
  scale_fill_brewer(palette = "Set3") +
  #  ylim(0, 500) +
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 20),  # Center title, bold, and larger
    axis.title.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels
  ) +
    geom_text(aes(label = Frequency), vjust = -0.5, size = 3, position = position_stack(vjust = 0.5))

  #geom_text(aes(label = paste0(round(Percentage, 1), "%")), 
  #          vjust = -0.5, size = 3, position = position_stack(vjust = 0.5))  # Add text labels inside bars

# filter Gov institutes only
sams_student_Gov_data = sams_student_data %>% filter(sams_student_data$TypeofInstitute == "Govt.")

# Calculate frequency for Gov. institutes only
freq_df_gov_gender <- sams_student_Gov_data %>%
  group_by(Year, Gender) %>%
  summarise(Frequency = n()) %>%
  mutate(Percentage = Frequency / sum(Frequency) * 100) %>%
  ungroup()

# Create the Gov only stacked bar chart
ggplot(freq_df_gov_gender, aes(x = as.factor(Year), y = Frequency, fill = Gender)) +
  geom_bar(stat = "identity", color = "black") +  # Bar color and border
  labs(
    title = "Count SAMS Reported by Year and Gender in Gov Institutes only",
    x = "Year",
    y = "Count",
    fill = "Gender"
  ) +
  scale_fill_brewer(palette = "Set3") +
  #  ylim(0, 500) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 20),  # Center title, bold, and larger
    axis.title.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels
  ) +
  geom_text(aes(label = Frequency), vjust = -0.5, size = 3, position = position_stack(vjust = 0.5))


# Difference between data count and SAMS website 
# -294 (2023-24)
# -3246 (2022-23)
# -13029 (2021-22)
# (2020-21)

```

# Count of Institutes - ITI, Poly & PDIS by Year


```{r}
# 
unique_institues = df %>%
  filter(Year == "2024" & module =="ITI") %>%  # Change "ITI" to "Diploma" or "PDIS" as needed
  summarise(Unique_Count = n_distinct(ReportedInstitute))

print(unique_institues)

```

# "Percentage" validation exercise
```{r}
# Convert SecuredMarks and TotalMarks to numeric
df <- df %>%
  mutate(
    SecuredMarks = as.numeric(as.character(SecuredMarks)),
    TotalMarks = as.numeric(as.character(TotalMarks)),
    Percentage = as.numeric(as.character(Percentage))
  )

# Calculate the expected percentage
df <- df %>%
  mutate(expected_percentage = (SecuredMarks/TotalMarks) * 100)

# Compare the expected percentage with the actual percentage
df <- df %>%
  mutate(is_correct = round(expected_percentage, 2) == round(Percentage, 2))

# Count the number of correct rows
num_correct <- sum(df$is_correct, na.rm = TRUE)

selected_columns = df %>%
  filter(is_correct == FALSE) %>%
  select(Percentage,expected_percentage)

hist(selected_columns$Percentage)
hist(selected_columns$expected_percentage)

print(paste("Number of correct rows:", num_correct))
```




## Visualizations 

# "Percentage" Distribution Visuals
```{r}
library(ggplot2)

# Subset to module 
df_ITI = df %>% filter(df$module == "ITI")
df_Diploma = df %>% filter(df$module == "Diploma")
df_PDIS = df %>% filter(df$module == "PDIS")


# Custom function to create the histogram
plot_percentage_histogram <- function(data, percentage_col) {
  # Create histogram
  p <- ggplot(data, aes_string(x = percentage_col)) +
    geom_histogram(
      aes(y = ..density.. * 100),  # Use density and multiply by 100 for percentage
      binwidth = 5, 
      fill = "#2E86C1", 
      color = "#1B4F72", 
      alpha = 0.7
    ) +
    labs(
      title = paste("Distribution of", percentage_col, "Scores"),
      x = "Percentage",
      y = "Percentage of Total"
    ) +
    scale_x_continuous(
      limits = c(0, 100),  # Limit the x-axis from 0 to 100
      breaks = seq(0, 100, by = 10)  # Set breaks at intervals of 10
    ) +
    scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentage
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
      axis.title = element_text(size = 12),
      axis.text = element_text(size = 10),
      panel.grid.major = element_line(color = "#D5D8DC"),
      panel.grid.minor = element_blank(),
      panel.background = element_rect(fill = "#F7F9F9")
    ) +
    geom_density(
      aes(y = ..density.. * 100),  # Adjust density line to match percentage scale
      color = "#CB4335", 
      fill = "#F5B7B1", 
      alpha = 0.3,
      adjust = 1
    )
  
  return(p)
}


# Plot percentage for ITI
histogram1 <- plot_percentage_histogram(df_ITI, "expected_percentage")
print(histogram1)

# Plot percentage for Diploma
histogram2 <- plot_percentage_histogram(df_Diploma, "expected_percentage")
print(histogram2)

# Plot percentage for Diploma
histogram3 <- plot_percentage_histogram(df_PDIS, "expected_percentage")
print(histogram3)

```

# Marks Percentage Histograms by Gender

Who are the ones applying and getting in ITI, Polytechnics and PDIS? The minimum marks percentage to apply for ITIs is 33% in 10th or 8th Grade (minimum passing marks) depending on the trade. To apply for Diploma through Fresh entry in Engineering or Beauty Culture courses, one has to pass 10th grade at 33%. For Diploma courses in Hotel Management & Catering Technology, Modern Office Management, and Pharmacy, applicant should be a 10+2 pass in science/commerce/arts. 
For Lateral Entry to Engineering course, one has to Pass in +2 Science or +2 Vocational in an Engineering Trade/ ITI Pass from 2 years ITI in Engineering Trade or COE Trade with pass in 10th. 

```{r}
plot_overlapping_histograms <- function(data, percentage_col, module_col) {
  
  # Create the histogram and density plot
  p <- ggplot(data, aes_string(x = percentage_col, fill = module_col, color = module_col)) +
    geom_histogram(
      aes(y = ..density.. * 100),  # Use density and multiply by 100 for percentage
      binwidth = 5,  # Adjust binwidth as needed
      position = "identity",  # Overlap histograms
      alpha = 0.5,  # Transparency of histograms
      color = "black"
    ) +
    geom_density(
      aes(y = ..density.. * 100),  # Adjust density line to match percentage scale
      adjust = 1,  # Smoothness of density line
      alpha = 0.5  # Transparency of density line
    ) +
    labs(
      title = "Distribution of Percentage Scores by Module",
      x = "Percentage",
      y = "Percentage of Total"
    ) +
    scale_x_continuous(
      limits = c(0, 100),  # Limit the x-axis from 0 to 100
      breaks = seq(0, 100, by = 10)  # Set breaks at intervals of 10
    ) +
    scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentage
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
      axis.title = element_text(size = 12),
      axis.text = element_text(size = 10),
      panel.grid.major = element_line(color = "#D5D8DC"),
      panel.grid.minor = element_blank(),
      panel.background = element_rect(fill = "#F7F9F9")
    )
  
  return(p)
}

# Assuming your dataset is named 'df_ITI' and has columns 'percentage_score' and 'Gender'
bar_plot <- plot_overlapping_histograms(df, "expected_percentage", "module")
print(bar_plot)


```
We can see from the above histogram, ITI applicants have a more right skew compared to Diploma applicants in their percentage marks. ITI applicants generally tend to be academically weaker students. However, it is not clear whether the reported percentage marks for Diploma applicants is 10th, 12th or ITI.
Now, let's compare the marks distribution between private and government institutes. 

# Private Vs Government - Percentage Marks

```{r}
plot_pvt_gov_overlapping_histograms <- function(data, percentage_col, type_col) {
  
  # Create the histogram and density plot
  p <- ggplot(data, aes_string(x = percentage_col, fill = type_col, color = type_col)) +
    geom_histogram(
      aes(y = ..density.. * 100),  # Use density and multiply by 100 for percentage
      binwidth = 5,  # Adjust binwidth as needed
      position = "identity",  # Overlap histograms
      alpha = 0.5,  # Transparency of histograms
      color = "black"
    ) +
    geom_density(
      aes(y = ..density.. * 100),  # Adjust density line to match percentage scale
      adjust = 1,  # Smoothness of density line
      alpha = 0.5  # Transparency of density line
    ) +
    labs(
      title = "Distribution of Percentage Scores by Institute Type",
      x = "Percentage",
      y = "Percentage of Total"
    ) +
    scale_x_continuous(
      limits = c(0, 100),  # Limit the x-axis from 0 to 100
      breaks = seq(0, 100, by = 10)  # Set breaks at intervals of 10
    ) +
    scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentage
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
      axis.title = element_text(size = 12),
      axis.text = element_text(size = 10),
      panel.grid.major = element_line(color = "#D5D8DC"),
      panel.grid.minor = element_blank(),
      panel.background = element_rect(fill = "#F7F9F9")
    )
  
  return(p)
}

# Assuming your dataset is named 'df_ITI' and has columns 'percentage_score' and 'TypeofInstitute'
bar_ITItype_plot <- plot_pvt_gov_overlapping_histograms(df_ITI, "expected_percentage", "TypeofInstitute")
print(bar_ITItype_plot)

bar_Diplomatype_plot <- plot_pvt_gov_overlapping_histograms(df_Diploma, "expected_percentage", "TypeofInstitute")
print(bar_Diplomatype_plot)



```
From the above graphs, applicants in Gov institutes tend to have higher percentage marks than private institutes across ITI and Diploma courses. 


# Gender wise - Percentage Marks

```{r}
# Exclude Transgender from dataset
excl_trans_df_ITI = df_ITI %>% filter(Gender != "Transgender")
excl_trans_df_Diploma = df_Diploma %>% filter(Gender != "Transgender")

plot_gender_overlapping_histograms <- function(data, percentage_col, gender_col) {
  
  # Create the histogram and density plot
  p <- ggplot(data, aes_string(x = percentage_col, fill = gender_col, color = gender_col)) +
    geom_histogram(
      aes(y = ..density.. * 100),  # Use density and multiply by 100 for percentage
      binwidth = 5,  # Adjust binwidth as needed
      position = "identity",  # Overlap histograms
      alpha = 0.5,  # Transparency of histograms
      color = "black"
    ) +
    geom_density(
      aes(y = ..density.. * 100),  # Adjust density line to match percentage scale
      adjust = 1,  # Smoothness of density line
      alpha = 0.5  # Transparency of density line
    ) +
    labs(
      title = "Distribution of Percentage Scores by Institute Type",
      x = "Percentage",
      y = "Percentage of Total"
    ) +
    scale_x_continuous(
      limits = c(0, 100),  # Limit the x-axis from 0 to 100
      breaks = seq(0, 100, by = 10)  # Set breaks at intervals of 10
    ) +
    scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentage
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
      axis.title = element_text(size = 12),
      axis.text = element_text(size = 10),
      panel.grid.major = element_line(color = "#D5D8DC"),
      panel.grid.minor = element_blank(),
      panel.background = element_rect(fill = "#F7F9F9")
    )
  
  return(p)
}

# Assuming your dataset is named 'df_ITI' and has columns 'percentage_score' and 'TypeofInstitute'
bar_ITItype_plot <- plot_pvt_gov_overlapping_histograms(excl_trans_df_ITI, "expected_percentage", "Gender")
print(bar_ITItype_plot)

bar_Diplomatype_plot <- plot_pvt_gov_overlapping_histograms(excl_trans_df_Diploma, "expected_percentage", "Gender")
print(bar_Diplomatype_plot)



```
There is very minimal disparity between two genders in ITIs, however the disparity is relatively higher in Diploma with Male being more skewed to the left than Female. 

# Percentage Time Trends

```{r}

# Custom function to create a bar plot of the mean percentage score by year
plot_mean_percentage_by_year <- function(data, year_col, percentage_col) {
  
  # Calculate the arithmetic mean of percentage_score by year
  mean_data <- data %>%
    group_by_at(year_col) %>%
    summarise(mean_percentage = mean(.data[[percentage_col]], na.rm = TRUE))
  
  # Create the bar plot
  p <- ggplot(mean_data, aes_string(x = year_col, y = "mean_percentage")) +
    geom_bar(stat = "identity", fill = "#3498DB", color = "#1F78B4", alpha = 0.8) +
    labs(
      title = paste("Mean Percentage Score by Year"),
      x = "Year",
      y = "Mean Percentage Score"
    ) +
    ylim(0,100) +
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
      axis.title = element_text(size = 14),
      axis.text = element_text(size = 12),
      panel.grid.major = element_line(color = "#D5D8DC"),
      panel.grid.minor = element_blank(),
      panel.background = element_rect(fill = "#F7F9F9"),
      axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels for better readability
    ) +
    geom_text(aes(label = round(mean_percentage, 2)), vjust = -0.5, size = 4)
  
  return(p)
}

# Example usage
# Assuming your dataset is named 'data' and has columns 'Year' and 'percentage_score'
bar_ITI_time_plot <- plot_mean_percentage_by_year(df_ITI, "Year", "expected_percentage")
print(bar_ITI_time_plot)

bar_Diploma_time_plot <- plot_mean_percentage_by_year(df_Diploma, "Year", "expected_percentage")
print(bar_Diploma_time_plot)

bar_PDIS_time_plot <- plot_mean_percentage_by_year(df_PDIS, "Year", "expected_percentage")
print(bar_PDIS_time_plot)

```
# Percentage Time Trends by Gender

```{r}

plot_mean_percentage_by_year_gender <- function(data, year_col, percentage_col, gender_col) {
# Calculate the arithmetic mean of percentage_score by year and gender
  mean_data <- data %>%
    group_by_at(c(year_col, gender_col)) %>%
    summarise(mean_percentage = mean(.data[[percentage_col]], na.rm = TRUE), .groups = 'drop')
  
 # Create the bar plot
  p <- ggplot(mean_data, aes_string(x = year_col, y = "mean_percentage", fill = gender_col)) +
    geom_bar(stat = "identity", position = position_dodge(), color = "#1F78B4", alpha = 0.8) +
    labs(
      title = "Mean Percentage Score by Year and Gender",
      x = "Year",
      y = "Mean Percentage Score"
    ) +
    scale_fill_manual(values = c("Male" = "#3498DB", "Female" = "#E74C3C")) +
    theme_minimal() +
    ylim(0,100) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
      axis.title = element_text(size = 14),
      axis.text = element_text(size = 12),
      panel.grid.major = element_line(color = "#D5D8DC"),
      panel.grid.minor = element_blank(),
      panel.background = element_rect(fill = "#F7F9F9"),
      axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels for better readability
    ) +
    geom_text(aes(label = round(mean_percentage, 2)), position = position_dodge(width = 0.9), vjust = -0.5, size = 3)
  
  return(p)
}

# Example usage
# Assuming your dataset is named 'data' and has columns 'year', 'percentage_score', and 'gender'
bar_plot_ITI <- plot_mean_percentage_by_year_gender(df_ITI, "Year", "expected_percentage","Gender")
print(bar_plot_ITI)

bar_plot_Diploma <- plot_mean_percentage_by_year_gender(df_Diploma, "Year", "expected_percentage","Gender")
print(bar_plot_Diploma)

bar_plot_PDIS <- plot_mean_percentage_by_year_gender(df_PDIS, "Year", "expected_percentage","Gender")
print(bar_plot_PDIS)

```




