---
title: "SAMS Student Data Diagnostics"
output: html_document
date: "2024-08-29"
---
This document provides a diagnostics of SAMS data - API access - CSM. 


# Load the data set
```{r}
sams_student_data <- read_csv("C:/Users/Admin/Box/Amitav/Skills/student_data.csv")


```


# Check for Duplicates in id variables
```{r}
# Check for the non-unique Barcode × Year. Generates a column which is a count of duplicates
df_duplicates_barcode <- sams_student_data %>%
  group_by(Barcode,Year) %>%
  summarize(num_duplicates = n())

 
#filtered_sams = sams_student_data %>% filter(Barcode %in% duplicate_barcodes$Barcode)
#view(filtered_sams)
#remove(sams_student_data)
#print(df_duplicates_barcode$Barcode>1)

#duplicate_barcodes <- df_duplicates_barcode %>%
 # filter(num_duplicates > 1) %>%   # Filter rows where col1 is greater than 1
  #select(Barcode)           # Select only col1

```

# Check if all the institutes in 2024 (ITI, Diploma and PDIS) are present in the dataset
This should be done after the admission cycle has ended to check if the "college participation" figures for 2024 on SAMS website matches with the SAMS dataset. 
```{r}
## For ITIs
missing_itis = setdiff(gsub(" ", "", iti_list_2024_25$Institute_Name), gsub(" ", "", sams_student_data$ReportedInstitute))

print("Number of ITIs missing from SAMS")
print(missing_itis)

remove(d_list,e_list)
## For Diploma

## For PDIS

```

# Year-wise count
```{r}
# Calculate frequency of Gender and Year for all institutes
freq_df_gender <- sams_student_data %>%
  group_by(Year, Gender) %>%
  summarise(Frequency = n()) %>%
  mutate(Percentage = Frequency / sum(Frequency) * 100) %>%
  ungroup()

# Create the stacked bar chart
ggplot(freq_df_gender, aes(x = as.factor(Year), y = Frequency, fill = Gender)) +
  geom_bar(stat = "identity", color = "black") +  # Bar color and border
  labs(
    title = "Count SAMS Reported by Year and Gender",
    x = "Year",
    y = "Count",
    fill = "Gender"
  ) +
  scale_fill_brewer(palette = "Set3") +
  #  ylim(0, 500) +
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 20),  # Center title, bold, and larger
    axis.title.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels
  ) +
    geom_text(aes(label = Frequency), vjust = -0.5, size = 3, position = position_stack(vjust = 0.5))

  #geom_text(aes(label = paste0(round(Percentage, 1), "%")), 
  #          vjust = -0.5, size = 3, position = position_stack(vjust = 0.5))  # Add text labels inside bars

# filter Gov institutes only
sams_student_Gov_data = sams_student_data %>% filter(sams_student_data$TypeofInstitute == "Govt.")

# Calculate frequency for Gov. institutes only
freq_df_gov_gender <- sams_student_Gov_data %>%
  group_by(Year, Gender) %>%
  summarise(Frequency = n()) %>%
  mutate(Percentage = Frequency / sum(Frequency) * 100) %>%
  ungroup()

# Create the Gov only stacked bar chart
ggplot(freq_df_gov_gender, aes(x = as.factor(Year), y = Frequency, fill = Gender)) +
  geom_bar(stat = "identity", color = "black") +  # Bar color and border
  labs(
    title = "Count SAMS Reported by Year and Gender in Gov Institutes only",
    x = "Year",
    y = "Count",
    fill = "Gender"
  ) +
  scale_fill_brewer(palette = "Set3") +
  #  ylim(0, 500) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 20),  # Center title, bold, and larger
    axis.title.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels
  ) +
  geom_text(aes(label = Frequency), vjust = -0.5, size = 3, position = position_stack(vjust = 0.5))


# Difference between data count and SAMS website 
# -294 (2023-24)
# -3246 (2022-23)
# -13029 (2021-22)
# (2020-21)

```



