---
title: "SAMS Student Data Diagnostics"
output: html_document
date: "2024-08-29"
---
This document provides a diagnostics of SAMS data - API access - CSM. 



# Import the .parquest datafile
```{r}
install.packages("arrow")
library(arrow)

# Replace 'your_file.parquet' with the path to your Parquet file
df <- read_parquet("C:/Users/Admin/Box/Amitav/Skills/student_data.parquet")

```



# Check for Duplicates in id variables
```{r}
# Check for the non-unique Barcode × Year. Generates a column which is a count of duplicates
df_duplicates_barcode <- df %>%
  group_by(Barcode) %>%
  summarize(num_duplicates = n())

duplicate_barcodes <- df_duplicates_barcode %>%
 filter(num_duplicates > 1) %>%   # Filter rows where num_duplicates are greater than 1
 select(Barcode)           # Select only those Barcodes


barcode_dupli_list = df %>% filter(Barcode %in% duplicate_barcodes$Barcode)
view(barcode_dupli_list)

write.xlsx(barcode_dupli_list, "barcode_duplicates_df.xlsx")

 
#filtered_sams = sams_student_data %>% filter(Barcode %in% duplicate_barcodes$Barcode)
#view(filtered_sams)
#remove(sams_student_data)
#print(df_duplicates_barcode$Barcode>1)

#duplicate_barcodes <- df_duplicates_barcode %>%
 # filter(num_duplicates > 1) %>%   # Filter rows where col1 is greater than 1
  #select(Barcode)           # Select only col1

```

# Check if all the institutes in 2024 (ITI, Diploma and PDIS) are present in the dataset
This should be done after the admission cycle has ended to check if the "college participation" figures for 2024 on SAMS website matches with the SAMS dataset. 
```{r}
## For ITIs
missing_itis = setdiff(gsub(" ", "", iti_list_2024_25$Institute_Name), gsub(" ", "", sams_student_data$ReportedInstitute))

print("Number of ITIs missing from SAMS")
print(missing_itis)

remove(d_list,e_list)
## For Diploma

## For PDIS

```



# Year-wise count
```{r}
# Calculate frequency of Gender and Year for all institutes
freq_df_gender <- df_ITI %>%
  group_by(Year, Gender) %>%
  summarise(Frequency = n()) %>%
  mutate(Percentage = Frequency / sum(Frequency) * 100) %>%
  ungroup()

# Create the stacked bar chart
ggplot(freq_df_gender, aes(x = as.factor(Year), y = Frequency, fill = Gender)) +
  geom_bar(stat = "identity", color = "black") +  # Bar color and border
  labs(
    title = "Count Students Reported by Gender in ITIs",
    x = "Year",
    y = "Count",
    fill = "Gender"
  ) +
  scale_fill_brewer(palette = "Set3") +
  #  ylim(0, 500) +
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 20),  # Center title, bold, and larger
    axis.title.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels
  ) +
    geom_text(aes(label = Frequency), vjust = -0.5, size = 3, position = position_stack(vjust = 0.5)) #+
 # geom_text(aes(label = paste0(round(Percentage, 1), "%")), 
#          vjust = -0.5, size = 3, position = position_stack(vjust = 0.5))  # Add text labels inside bars

# filter Gov institutes only
df_ITI_gov = df %>% filter(df$TypeofInstitute == "Govt.")

# Calculate frequency for Gov. institutes only
freq_df_gov_gender <- df_ITI_gov %>%
  group_by(Year, Gender) %>%
  summarise(Frequency = n()) %>%
  mutate(Percentage = Frequency / sum(Frequency) * 100) %>%
  ungroup()

# Create the Gov only stacked bar chart
ggplot(freq_df_gov_gender, aes(x = as.factor(Year), y = Frequency, fill = Gender)) +
  geom_bar(stat = "identity", color = "black") +  # Bar color and border
  labs(
    title = "Count Students Reported by Gender in Gov ITIs only",
    x = "Year",
    y = "Count",
    fill = "Gender"
  ) +
  scale_fill_brewer(palette = "Set3") +
  ylim(0, 30000) +
  theme_minimal(base_size = 10) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 20),  # Center title, bold, and larger
    axis.title.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels
  ) +
#  geom_text(aes(label = Frequency), vjust = -0.5, size = 3, position = position_stack(vjust = 0.5))
 geom_text(aes(label = paste0(round(Percentage, 1), "%")), 
          vjust = -0.5, size = 3, position = position_stack(vjust = 0.5))  # Add text labels inside bars

# Difference between data count and SAMS website 
# -294 (2023-24)
# -3246 (2022-23)
# -13029 (2021-22)
# (2020-21)

```

# Count of Institutes - ITI, Poly & PDIS by Year


```{r}
# 
unique_institues = df %>%
  filter(Year == "2024" & module =="ITI") %>%  # Change "ITI" to "Diploma" or "PDIS" as needed
  summarise(Unique_Count = n_distinct(ReportedInstitute))

print(unique_institues)

```

# "Percentage" validation exercise
```{r}
# Convert SecuredMarks and TotalMarks to numeric
df <- df %>%
  mutate(
    SecuredMarks = as.numeric(as.character(SecuredMarks)),
    TotalMarks = as.numeric(as.character(TotalMarks)),
    Percentage = as.numeric(as.character(Percentage))
  )

# Calculate the expected percentage
df <- df %>%
  mutate(expected_percentage = (SecuredMarks/TotalMarks) * 100)

# Compare the expected percentage with the actual percentage
df <- df %>%
  mutate(is_correct = round(expected_percentage, 2) == round(Percentage, 2))

# Count the number of correct rows
num_correct <- sum(df$is_correct, na.rm = TRUE)

selected_columns = df %>%
  filter(is_correct == FALSE) %>%
  select(Percentage,expected_percentage)

hist(selected_columns$Percentage)
hist(selected_columns$expected_percentage)

print(paste("Number of correct rows:", num_correct))
```




## Visualizations 

# "Percentage" Distribution Visuals
```{r}
library(ggplot2)

# Subset to module 
df_ITI = df %>% filter(df$module == "ITI")
df_Diploma = df %>% filter(df$module == "Diploma")
df_PDIS = df %>% filter(df$module == "PDIS")


# Custom function to create the histogram
plot_percentage_histogram <- function(data, percentage_col) {
  # Create histogram
  p <- ggplot(data, aes_string(x = percentage_col)) +
    geom_histogram(
      aes(y = ..density.. * 100),  # Use density and multiply by 100 for percentage
      binwidth = 5, 
      fill = "#2E86C1", 
      color = "#1B4F72", 
      alpha = 0.7
    ) +
    labs(
      title = paste("Distribution of", percentage_col, "Scores"),
      x = "Percentage",
      y = "Percentage of Total"
    ) +
    scale_x_continuous(
      limits = c(0, 100),  # Limit the x-axis from 0 to 100
      breaks = seq(0, 100, by = 10)  # Set breaks at intervals of 10
    ) +
    scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentage
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
      axis.title = element_text(size = 12),
      axis.text = element_text(size = 10),
      panel.grid.major = element_line(color = "#D5D8DC"),
      panel.grid.minor = element_blank(),
      panel.background = element_rect(fill = "#F7F9F9")
    ) +
    geom_density(
      aes(y = ..density.. * 100),  # Adjust density line to match percentage scale
      color = "#CB4335", 
      fill = "#F5B7B1", 
      alpha = 0.3,
      adjust = 1
    )
  
  return(p)
}


# Plot percentage for ITI
histogram1 <- plot_percentage_histogram(df_ITI, "expected_percentage")
print(histogram1)

# Plot percentage for Diploma
histogram2 <- plot_percentage_histogram(df_Diploma, "expected_percentage")
print(histogram2)

# Plot percentage for Diploma
histogram3 <- plot_percentage_histogram(df_PDIS, "expected_percentage")
print(histogram3)

```

# Marks Percentage Histograms by Gender

Who are the ones applying and getting in ITI, Polytechnics and PDIS? The minimum marks percentage to apply for ITIs is 33% in 10th or 8th Grade (minimum passing marks) depending on the trade. To apply for Diploma through Fresh entry in Engineering or Beauty Culture courses, one has to pass 10th grade at 33%. For Diploma courses in Hotel Management & Catering Technology, Modern Office Management, and Pharmacy, applicant should be a 10+2 pass in science/commerce/arts. 
For Lateral Entry to Engineering course, one has to Pass in +2 Science or +2 Vocational in an Engineering Trade/ ITI Pass from 2 years ITI in Engineering Trade or COE Trade with pass in 10th. 

```{r}
plot_overlapping_histograms <- function(data, percentage_col, module_col) {
  
  # Create the histogram and density plot
  p <- ggplot(data, aes_string(x = percentage_col, fill = module_col, color = module_col)) +
    geom_histogram(
      aes(y = ..density.. * 100),  # Use density and multiply by 100 for percentage
      binwidth = 5,  # Adjust binwidth as needed
      position = "identity",  # Overlap histograms
      alpha = 0.5,  # Transparency of histograms
      color = "black"
    ) +
    geom_density(
      aes(y = ..density.. * 100),  # Adjust density line to match percentage scale
      adjust = 1,  # Smoothness of density line
      alpha = 0.5  # Transparency of density line
    ) +
    labs(
      title = "Distribution of Percentage Scores by Module",
      x = "Percentage",
      y = "Percentage of Total"
    ) +
    scale_x_continuous(
      limits = c(0, 100),  # Limit the x-axis from 0 to 100
      breaks = seq(0, 100, by = 10)  # Set breaks at intervals of 10
    ) +
    scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentage
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
      axis.title = element_text(size = 12),
      axis.text = element_text(size = 10),
      panel.grid.major = element_line(color = "#D5D8DC"),
      panel.grid.minor = element_blank(),
      panel.background = element_rect(fill = "#F7F9F9")
    )
  
  return(p)
}

# Assuming your dataset is named 'df_ITI' and has columns 'percentage_score' and 'Gender'
bar_plot <- plot_overlapping_histograms(df, "expected_percentage", "module")
print(bar_plot)


```
We can see from the above histogram, ITI applicants have a more right skew compared to Diploma applicants in their percentage marks. ITI applicants generally tend to be academically weaker students. However, it is not clear whether the reported percentage marks for Diploma applicants is 10th, 12th or ITI.
Now, let's compare the marks distribution between private and government institutes. 

# Private Vs Government - Percentage Marks

```{r}
plot_pvt_gov_overlapping_histograms <- function(data, percentage_col, type_col) {
  
  # Create the histogram and density plot
  p <- ggplot(data, aes_string(x = percentage_col, fill = type_col, color = type_col)) +
    geom_histogram(
      aes(y = ..density.. * 100),  # Use density and multiply by 100 for percentage
      binwidth = 5,  # Adjust binwidth as needed
      position = "identity",  # Overlap histograms
      alpha = 0.5,  # Transparency of histograms
      color = "black"
    ) +
    geom_density(
      aes(y = ..density.. * 100),  # Adjust density line to match percentage scale
      adjust = 1,  # Smoothness of density line
      alpha = 0.5  # Transparency of density line
    ) +
    labs(
      title = "Distribution of Percentage Scores by Institute Type",
      x = "Percentage",
      y = "Percentage of Total"
    ) +
    scale_x_continuous(
      limits = c(0, 100),  # Limit the x-axis from 0 to 100
      breaks = seq(0, 100, by = 10)  # Set breaks at intervals of 10
    ) +
    scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentage
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
      axis.title = element_text(size = 12),
      axis.text = element_text(size = 10),
      panel.grid.major = element_line(color = "#D5D8DC"),
      panel.grid.minor = element_blank(),
      panel.background = element_rect(fill = "#F7F9F9")
    )
  
  return(p)
}

# Assuming your dataset is named 'df_ITI' and has columns 'percentage_score' and 'TypeofInstitute'
bar_ITItype_plot <- plot_pvt_gov_overlapping_histograms(df_ITI, "expected_percentage", "TypeofInstitute")
print(bar_ITItype_plot)

bar_Diplomatype_plot <- plot_pvt_gov_overlapping_histograms(df_Diploma, "expected_percentage", "TypeofInstitute")
print(bar_Diplomatype_plot)



```
From the above graphs, applicants in Gov institutes tend to have higher percentage marks than private institutes across ITI and Diploma courses. 


# Gender wise - Percentage Marks

```{r}
# Exclude Transgender from dataset
excl_trans_df_ITI = df_ITI %>% filter(Gender != "Transgender")
excl_trans_df_Diploma = df_Diploma %>% filter(Gender != "Transgender")

plot_gender_overlapping_histograms <- function(data, percentage_col, gender_col) {
  
  # Create the histogram and density plot
  p <- ggplot(data, aes_string(x = percentage_col, fill = gender_col, color = gender_col)) +
    geom_histogram(
      aes(y = ..density.. * 100),  # Use density and multiply by 100 for percentage
      binwidth = 5,  # Adjust binwidth as needed
      position = "identity",  # Overlap histograms
      alpha = 0.5,  # Transparency of histograms
      color = "black"
    ) +
    geom_density(
      aes(y = ..density.. * 100),  # Adjust density line to match percentage scale
      adjust = 1,  # Smoothness of density line
      alpha = 0.5  # Transparency of density line
    ) +
    labs(
      x = "Marks Percentage",
      y = "Percentage of Total"
    ) +
    scale_x_continuous(
      limits = c(0, 100),  # Limit the x-axis from 0 to 100
      breaks = seq(0, 100, by = 10)  # Set breaks at intervals of 10
    ) +
    scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Format y-axis as percentage
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
      axis.title = element_text(size = 12),
      axis.text = element_text(size = 10),
      panel.grid.major = element_line(color = "#D5D8DC"),
      panel.grid.minor = element_blank(),
      panel.background = element_rect(fill = "#F7F9F9")
    )
  
  return(p)
}

# Assuming your dataset is named 'df_ITI' and has columns 'percentage_score' and 'TypeofInstitute'
bar_ITItype_plot <- plot_pvt_gov_overlapping_histograms(excl_trans_df_ITI, "expected_percentage", "Gender")
print(bar_ITItype_plot)

bar_Diplomatype_plot <- plot_pvt_gov_overlapping_histograms(excl_trans_df_Diploma, "expected_percentage", "Gender")
print(bar_Diplomatype_plot)



```
There is very minimal disparity between two genders in ITIs, however the disparity is relatively higher in Diploma with Male being more skewed to the left than Female. 

# Percentage Time Trends

```{r}

# Custom function to create a bar plot of the mean percentage score by year
plot_mean_percentage_by_year <- function(data, year_col, percentage_col) {
  
  # Calculate the arithmetic mean of percentage_score by year
  mean_data <- data %>%
    group_by_at(year_col) %>%
    summarise(mean_percentage = mean(.data[[percentage_col]], na.rm = TRUE))
  
  # Create the bar plot
  p <- ggplot(mean_data, aes_string(x = year_col, y = "mean_percentage")) +
    geom_bar(stat = "identity", fill = "#3498DB", color = "#1F78B4", alpha = 0.8) +
    labs(
      title = paste("Mean Percentage Score by Year"),
      x = "Year",
      y = "Mean Percentage Score"
    ) +
    ylim(0,100) +
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
      axis.title = element_text(size = 14),
      axis.text = element_text(size = 12),
      panel.grid.major = element_line(color = "#D5D8DC"),
      panel.grid.minor = element_blank(),
      panel.background = element_rect(fill = "#F7F9F9"),
      axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels for better readability
    ) +
    geom_text(aes(label = round(mean_percentage, 2)), vjust = -0.5, size = 4)
  
  return(p)
}

# Example usage
# Assuming your dataset is named 'data' and has columns 'Year' and 'percentage_score'
bar_ITI_time_plot <- plot_mean_percentage_by_year(df_ITI, "Year", "expected_percentage")
print(bar_ITI_time_plot)

bar_Diploma_time_plot <- plot_mean_percentage_by_year(df_Diploma, "Year", "expected_percentage")
print(bar_Diploma_time_plot)

bar_PDIS_time_plot <- plot_mean_percentage_by_year(df_PDIS, "Year", "expected_percentage")
print(bar_PDIS_time_plot)

```

# Percentage Time Trends by Gender

```{r}

plot_mean_percentage_by_year_gender <- function(data, year_col, percentage_col, gender_col) {
# Calculate the arithmetic mean of percentage_score by year and gender
  mean_data <- data %>%
    group_by_at(c(year_col, gender_col)) %>%
    summarise(mean_percentage = mean(.data[[percentage_col]], na.rm = TRUE), .groups = 'drop')
  
 # Create the bar plot
  p <- ggplot(mean_data, aes_string(x = year_col, y = "mean_percentage", fill = gender_col)) +
    geom_bar(stat = "identity", position = position_dodge(), color = "#1F78B4", alpha = 0.8) +
    labs(
      title = "Mean Percentage Score by Year and Gender",
      x = "Year",
      y = "Mean Percentage Score"
    ) +
    scale_fill_manual(values = c("Male" = "#3498DB", "Female" = "#E74C3C")) +
    theme_minimal() +
    ylim(0,100) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
      axis.title = element_text(size = 14),
      axis.text = element_text(size = 12),
      panel.grid.major = element_line(color = "#D5D8DC"),
      panel.grid.minor = element_blank(),
      panel.background = element_rect(fill = "#F7F9F9"),
      axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels for better readability
    ) +
    geom_text(aes(label = round(mean_percentage, 2)), position = position_dodge(width = 0.9), vjust = -0.5, size = 3)
  
  return(p)
}

# Example usage
# Assuming your dataset is named 'data' and has columns 'year', 'percentage_score', and 'gender'
bar_plot_ITI <- plot_mean_percentage_by_year_gender(df_ITI, "Year", "expected_percentage","Gender")
print(bar_plot_ITI)

bar_plot_Diploma <- plot_mean_percentage_by_year_gender(df_Diploma, "Year", "expected_percentage","Gender")
print(bar_plot_Diploma)

bar_plot_PDIS <- plot_mean_percentage_by_year_gender(df_PDIS, "Year", "expected_percentage","Gender")
print(bar_plot_PDIS)

```




# Income Male vs Female (Pvt. vs Public)
```{r}

# Calculate total counts for each gender for Gov
df_ITI_summary <- df_ITI_gov %>%
  group_by(Gender, AnnualIncome) %>%
  summarise(count = n(), .groups = 'drop') %>%
  group_by(Gender) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(df_ITI_summary, aes(x = AnnualIncome, y = percentage, fill = Gender)) +
  geom_bar(stat = "identity", position = "dodge", color = "white") +
  geom_text(aes(label = sprintf("%.1f%%", percentage)), 
            position = position_dodge(width = 0.9), 
            vjust = -0.5, size = 2) +
  labs(
    title = "Percentage Distribution of Income by Gender in Gov ITIs",
    x = "Annual Household Income (Rs.)",
    y = "Percentage",
    fill = "Gender"
  ) +
  ylim(0,100) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.title = element_text(size = 12)
  )

# Filter only private ITIs
df_ITI_pvt = df_ITI %>% filter(df_ITI$TypeofInstitute == "Pvt.")

# Calculate total counts for each gender for Gov
df_ITI_summary <- df_ITI_pvt %>%
  group_by(Gender, AnnualIncome) %>%
  summarise(count = n(), .groups = 'drop') %>%
  group_by(Gender) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(df_ITI_summary, aes(x = AnnualIncome, y = percentage, fill = Gender)) +
  geom_bar(stat = "identity", position = "dodge", color = "white") +
  geom_text(aes(label = sprintf("%.1f%%", percentage)), 
            position = position_dodge(width = 0.9), 
            vjust = -0.5, size = 2) +
  labs(
    title = "Percentage Distribution of Income by Gender in Pvt ITIs",
    x = "Annual Household Income (Rs.)",
    y = "Percentage",
    fill = "Gender"
  ) +
  ylim(0,100) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.title = element_text(size = 12)
  )






```

# Religion Male vs Female

```{r}
# Calculate total counts for each gender
df_ITI_summary <- df_ITI %>%
  group_by(Gender, ReligionName) %>%
  summarise(count = n(), .groups = 'drop') %>%
  group_by(Gender) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(df_ITI_summary, aes(x = ReligionName, y = percentage, fill = Gender)) +
  geom_bar(stat = "identity", position = "dodge", color = "white") +
  geom_text(aes(label = sprintf("%.1f%%", percentage)), 
            position = position_dodge(width = 0.9), 
            vjust = -0.5, size = 2) +
  labs(
    title = "Percentage Distribution of Religion by Gender in ITIs",
    x = "Religion",
    y = "Percentage",
    fill = "Gender"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.title = element_text(size = 12)
  )


```

# Social Category by Gender
```{r}
table(df_ITI_pvt$TypeofInstitute,df_ITI_pvt$module)
# Calculate total counts for each gender
df_ITI_summary <- df_ITI_gov %>%
  group_by(Gender, SocialCategory) %>%
  summarise(count = n(), .groups = 'drop') %>%
  group_by(Gender) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(df_ITI_summary, aes(x = SocialCategory, y = percentage, fill = Gender)) +
  geom_bar(stat = "identity", position = "dodge", color = "white") +
  geom_text(aes(label = sprintf("%.1f%%", percentage)), 
            position = position_dodge(width = 0.9), 
            vjust = -0.5, size = 3) +
  labs(
    title = "Social Category Distribution by Gender in Gov ITIs",
    x = "Social Category",
    y = "Percentage",
    fill = "Gender"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.title = element_text(size = 12)
  )

# Calculate total counts for each gender
df_ITI_summary <- df_ITI_pvt %>%
  group_by(Gender, SocialCategory) %>%
  summarise(count = n(), .groups = 'drop') %>%
  group_by(Gender) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(df_ITI_summary, aes(x = SocialCategory, y = percentage, fill = Gender)) +
  geom_bar(stat = "identity", position = "dodge", color = "white") +
  geom_text(aes(label = sprintf("%.1f%%", percentage)), 
            position = position_dodge(width = 0.9), 
            vjust = -0.5, size = 3) +
  labs(
    title = "Social Category Distribution by Gender in Pvt ITIs",
    x = "Social Category",
    y = "Percentage",
    fill = "Gender"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.title = element_text(size = 12)
  )

```

# Highest Qualification by Gender (Column needs cleaning)
```{r}

# Calculate total counts for each gender
df_ITI_summary <- df_ITI %>%
  group_by(Gender, HighestQualification) %>%
  summarise(count = n(), .groups = 'drop') %>%
  group_by(Gender) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(df_ITI_summary, aes(x = HighestQualification, y = percentage, fill = Gender)) +
  geom_bar(stat = "identity", position = "dodge", color = "white") +
  geom_text(aes(label = sprintf("%.1f%%", percentage)), 
            position = position_dodge(width = 0.9), 
            vjust = -0.5, size = 3) +
  labs(
    title = "Highest Qualification Distribution by Gender",
    x = "Highest Qualification",
    y = "Percentage",
    fill = "Gender"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.title = element_text(size = 12)
  )
```


# Course Period by Gender
```{r}
# Calculate total counts for each gender
df_ITI_summary <- df_ITI %>%
  group_by(Gender, CoursePeriod) %>%
  summarise(count = n(), .groups = 'drop') %>%
  group_by(Gender) %>%
  mutate(percentage = count / sum(count) * 100)

ggplot(df_ITI_summary, aes(x = CoursePeriod, y = percentage, fill = Gender)) +
  geom_bar(stat = "identity", position = "dodge", color = "white") +
  geom_text(aes(label = sprintf("%.1f%%", percentage)), 
            position = position_dodge(width = 0.9), 
            vjust = -0.5, size = 3) +
  labs(
    title = "ITI Course Period Distribution by Gender",
    x = "Course Period",
    y = "Percentage",
    fill = "Gender"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.title = element_text(size = 12)
  )
```


# Migration Analysis for ____

Top In-migration districts
```{r}
# Count the number of students by their home district and institute district
district_flow <- df_Diploma %>%
  group_by(District, InstituteDistrict) %>%
  summarise(student_count = n(), .groups = 'drop')


# Filter to include only students who study outside their home district
students_outside_home <- district_flow %>%
  filter(District != InstituteDistrict)

# Summarize the number of incoming students for each institute district
incoming_students <- students_outside_home %>%
  group_by(InstituteDistrict) %>%
  summarise(total_incoming_students = sum(student_count), .groups = 'drop') %>%
  arrange(desc(total_incoming_students))

# View the top institute districts by incoming students
top_institute_districts <- incoming_students %>%
  top_n(5, wt = total_incoming_students)  # Adjust the number to view more or fewer top districts

print(top_institute_districts)

# Bar plot to visualize the top institute districts
ggplot(top_institute_districts, aes(x = reorder(InstituteDistrict, -total_incoming_students), y = total_incoming_students)) +
  geom_bar(stat = "identity", fill = "#3498DB", alpha = 0.8) +
  labs(
    title = "Top Institute Districts by Incoming Students",
    x = "Institute District",
    y = "Total Incoming Students"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )

```

Top Out-migration districts
```{r}
district_percentage <- df_Diploma %>%
  group_by(District) %>%
  summarise(
    total_students = n(),
    students_outside = sum(District != InstituteDistrict),
    percentage_outside = (students_outside / total_students) * 100,
    .groups = 'drop'
  )

library(ggplot2)

ggplot(district_percentage, aes(x = reorder(District, percentage_outside), y = percentage_outside)) +
  geom_bar(stat = "identity", fill = "#3498DB", alpha = 0.8) +
  labs(
    title = "Percentage of Students Studying Outside Their Home District",
    x = "Home District",
    y = "Percentage of Students"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 10),
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.title = element_text(size = 12)
  ) +
  coord_flip()

```


# Net Migration Bar Chart
```{r}
df_female_Diploma = df_Diploma %>% filter(Gender == "Female")
# Calculate outgoing students (students leaving their home district)
outgoing_students <- df_female_Diploma %>%
  filter(District != InstituteDistrict) %>%
  group_by(District) %>%
  summarise(total_outgoing = n(), .groups = 'drop')

# Calculate incoming students (students entering the district from another district)
incoming_students <- df_female_Diploma %>%
  filter(District != InstituteDistrict) %>%
  group_by(InstituteDistrict) %>%
  summarise(total_incoming = n(), .groups = 'drop')

# Combine outgoing and incoming data
net_migration <- outgoing_students %>%
  full_join(incoming_students, by = c("District" = "InstituteDistrict")) %>%
  mutate(
    total_outgoing = replace_na(total_outgoing, 0),
    total_incoming = replace_na(total_incoming, 0),
    net_outmigration = total_outgoing - total_incoming
  )


library(ggpubr)

# Prepare the data for plotting
net_migration <- net_migration %>%
  mutate(migration_status = if_else(net_outmigration > 0, "Positive", "Negative"))
view(net_migration)
# Create the horizontal bar chart
ggbarplot(net_migration, x = "District", y = "net_outmigration",
          fill = "migration_status",  # Color bars based on migration status
          color = "white",            # Set bar border colors to white
          palette = c("Positive" = "yellow", "Negative" = "blue"),  # Define custom colors
          sort.val = "desc",          # Sort values in descending order
          sort.by.groups = FALSE,     # Don't sort inside each group
          x.text.angle = 90,          # Rotate x-axis labels
          ylab = "Net Outmigration",
          legend.title = "Migration Status",
          rotate = TRUE,              # Rotate the plot to make it horizontal
          ggtheme = theme_minimal()
          ) + scale_y_continuous(breaks = get_breaks(n = 7))

```


## Spatial Visualizations 

```{r}
## Student - Institute District Match
# Pvt. vs Gov.

table(sams_student_data$State)
filtered_sams_student = sams_student_data %>% filter(State == "Others")


```


# Check the big 5-10 ITIs by enrollment (by trade also) - boys/girls 

```{r}

# Step 1: Filter the data for the year 2023 and calculate counts
df_summary <- df_ITI %>%
  filter(Year == 2023) %>%
  group_by(Gender, ReportedInstitute) %>%
  summarise(student_count = n(), .groups = 'drop')

# Step 2: Select the top 10 institutes with the highest student counts
top_10_institutes <- df_summary %>%
  group_by(ReportedInstitute) %>%
  summarise(total_count = sum(student_count)) %>%
  slice_max(total_count, n = 10) %>%
  inner_join(df_summary, by = "ReportedInstitute")

# Step 3: Create the horizontal bar chart
ggplot(top_10_institutes, aes(x = reorder(ReportedInstitute, student_count), y = student_count, fill = Gender)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(
    title = "Number of Students by Gender in 2023 (Top 10 Institutes)",
    x = "Reported Institute",
    y = "Number of Students",
    fill = "Gender"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 12),
    axis.title = element_text(size = 12)
  )

## Top 10 institutes by girl students
# Step 1: Filter the data for the year 2023 and female students only
df_female_2023 <- df_ITI %>%
  filter(Year == 2023 & Gender == "Female")

# Step 2: Group by ReportedInstitute and summarize the count of female students
df_summary_female <- df_female_2023 %>%
  group_by(ReportedInstitute) %>%
  summarise(female_student_count = n(), .groups = 'drop')

# Step 3: Select the top 10 institutes with the highest female student counts
top_10_institutes_female <- df_summary_female %>%
  slice_max(female_student_count, n = 10)

# Step 4: Create the horizontal bar chart
ggplot(top_10_institutes_female, aes(x = reorder(ReportedInstitute, female_student_count), y = female_student_count, fill = ReportedInstitute)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  coord_flip() +
  labs(
    title = "Number of Female Students by Institute in 2023 (Top 10 Institutes)",
    x = "Reported Institute",
    y = "Number of Female Students"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 11),
    axis.title = element_text(size = 12),
    legend.position = "none"
  )

```
The districts in which the 10 institutes (total students) are situated are - 
Ganjam, Cuttack, Balasore, Khurda, Mayurbhanj and Keonjhar (East Odisha);
Sundergarh & Sambalpur (West Odisha).  

The districts in which the 10 institutes (total female students) are situated are - 
Ganjam, Cuttack, Balasore, Khurda, and Keonjhar (East Odisha);
Sundergarh, Sambalpur, Koraput (West Odisha).  
Angul (Central)

# Female Percentage in all Institutes over time
```{r}
# Step 1: Filter for government institutes
df_ITI_gov <- df_ITI %>%
  filter(TypeofInstitute == "Govt.")

# Step 2: Summarize the data (total students, female students, and female percentage) over time
df_summary_gov <- df_ITI_gov %>%
  filter(Year != "2024") %>%
  group_by(ReportedInstitute, Year) %>%
  summarise(
    total_students = n(),
    female_students = sum(Gender == "Female"),
    female_percentage = (female_students / total_students) * 100,
    .groups = 'drop'
  )

# Get the female_percentage for institutes in 2023
df_2023 <- df_summary_gov %>%
  filter(Year == 2023) %>%
  select(ReportedInstitute, female_percentage)

# Categorize institutes based on their 2023 female_percentage
df_2023 <- df_2023 %>%
  mutate(
    female_30_category_2023 = ifelse(female_percentage >= 30, "30% or more", "Below 30%"),
    female_sub_category_2023 = case_when(
      female_percentage > 0 & female_percentage <= 20 ~ "Between 0 to 20",
      female_percentage > 20 & female_percentage <= 30 ~ "Between 20 to 30",
      female_percentage > 30 & female_percentage <= 40 ~ "Between 30 to 40",
      female_percentage > 40 & female_percentage <= 50 ~ "Between 40 to 50",
      female_percentage > 50 ~ "Between 50 to 60"
    )
  )
  



# Step 4: Join this categorization back to the main summary dataframe
df_summary_gov <- df_summary_gov %>%
  left_join(df_2023 %>% select(ReportedInstitute, female_category_2023, female_sub_category_2023), by = "ReportedInstitute")


# Filter data for the specific year
df_sum_gov_2023 <- df_summary_gov %>%
  filter(Year == 2023)

view(df_sum_gov_2023)

top_10_female_percent <- df_sum_gov_2023 %>%
  slice_max(female_percentage, n = 15)

# Create the horizontal bar chart
ggplot(top_10_female_percent, aes(x = reorder(ReportedInstitute, female_percentage), y = female_percentage, fill = ReportedInstitute)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  coord_flip() +
  labs(
    title = "Female Percentage by Gov ITI Institutes in 2023 (Top 15 Institutes)",
    x = "Reported Institute",
    y = "Female Students (%)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 11),
    axis.title = element_text(size = 12),
    legend.position = "none"
  )

# Bar Plot
ggplot(df_sum_gov_2023, aes(x = reorder(ReportedInstitute, female_percentage), y = female_percentage, fill = female_sub_category_2023)) +
  geom_bar(stat = "identity") +
  coord_flip() +  # Flip coordinates to make it horizontal
  labs(
    title = "Female Percentage by Institute for 2023",
    x = "Institute",
    y = "Female Percentage",
    fill = "Female Percentage Category"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 8),
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.title = element_text(size = 12)
  )


# Bar Plot
df_count_2023 <- df_2023 %>%
  group_by(female_sub_category_2023) %>%
  summarise(count = n(), .groups = 'drop')
ggplot(df_count_2023, aes(x = female_sub_category_2023, y=count )) +
  geom_bar(stat="identity", fill="skyblue",  color = "black") +
  labs(title = "Distribution of Gov. ITIs by Female Percentage Category for 2023", x = "Female Percentage", y = "Number of ITIs", fill="Female Percentage Category") +
  ylim(0,35) +
  theme_minimal(base_size = 13) +  # Base text size
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),  # Center title, bold, and larger
    axis.title.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels
    plot.caption = element_text(size = 10)  # Increase bottom margin
  ) +
  geom_text(aes(label = count), vjust = -0.5, size = 4)  # Add text labels on top of bars

  

ggplot(data=df_summary_gov, aes(x=Year, y=female_percentage, group = ReportedInstitute, 
                                            colour = female_category_2023)) + 
  geom_line() +
  labs(y= "Female (%)", x = "Year") + 
  ggtitle("Percentage of Female Over Years in Top Institutes") + 
  geom_point()
 




```



# Female Percentage in Top 5 Institutes (in 2023) over time
```{r}

# Step 1: Filter for government institutes
df_ITI_gov <- df_ITI %>%
  filter(TypeofInstitute == "Govt.")

# Step 2: Summarize the data (total students, female students, and female percentage) over time
df_summary_gov <- df_ITI_gov %>%
  filter(Year != "2024") %>%
  group_by(ReportedInstitute, Year) %>%
  summarise(
    total_students = n(),
    female_students = sum(Gender == "Female"),
    female_percentage = (female_students / total_students) * 100,
    .groups = 'drop'
  )

# Step 3: Select the top 5 government institutes in 2023
top_5_institutes_2023 <- df_summary_gov %>%
  filter(Year == 2023) %>%
  slice_max(total_students, n = 5) %>%
  pull(ReportedInstitute)  # Extract the names of the top 5 institutes in 2023

# Step 4: Filter the original summary for only the top 5 institutes (over time)
df_top_5_summary <- df_summary_gov %>%
  filter(ReportedInstitute %in% top_5_institutes_2023)

ggplot(data=df_top_5_summary, aes(x=Year, y=female_percentage, group = ReportedInstitute, 
                                            colour = ReportedInstitute)) + 
  geom_line() +
  labs(y= "Female (%)", x = "Year") + 
  ggtitle("Percentage of Female Over Years in Top Gov ITIs") + 
  geom_point()
 




```



# Core Trades (identified by ChatGPT)
```{r}
# print list of trades in ITI
unique(df_ITI$ReportedBranchORTrade)

# "Core trades" identified by ChatGPT
core_trades <- c(
  "Electrician (NSQF)",
  "Fitter (NSQF)",
  "Mechanic (Motor Vehicle) (NSQF)",
  "Welder (NSQF)",
  "Electronics Mechanic (NSQF)",
  "Mechanic Diesel (NSQF)",
  "Turner (NSQF)",
  "Machinist (NSQF)",
  "Refrigeration & Air Conditioning Technician(NSQF)",
  "Information Technology(NSQF)"
)

filtered_data <- df_ITI %>%
  filter(ReportedBranchORTrade %in% core_trades)

# Calculate gender ratio by year and trade
gender_ratio <- filtered_data %>%
  group_by(Year, ReportedBranchORTrade, Gender) %>%
  summarise(Count = n()) %>%
  spread(key = Gender, value = Count, fill = 0) %>%
  mutate(Total = Male + Female, PercentFemale = (Female / Total) * 100)


# Create the sophisticated plot
ggplot(gender_ratio, aes(x = Year, y = PercentFemale, colour = ReportedBranchORTrade, group = ReportedBranchORTrade)) +
  geom_line() +
  geom_point(size = 1, shape = 21, fill = "white") +
  labs(
    title = "Percentage of Female Over Time by Core Trades",
    subtitle = "Gender distribution in core trades in Odisha ITIs (Female %)",
    x = "Year",
    y = "Female (%)",
    color = "Trade",
    linetype = "Trade"
  ) 

```


# Top Trades by various dimensions 
```{r}

gender_ratio <- df_ITI %>%
  group_by(Year, ReportedBranchORTrade, Gender) %>%
  summarise(Count = n()) %>%
  spread(key = Gender, value = Count, fill = 0) %>%
  mutate(Total = Male + Female, PercentFemale = (Female / Total) * 100) # Calculate ratio as Female Share (%)
view(gender_ratio)

gender_ratio_excl_2024 = gender_ratio %>% filter(Year != "2024")
view(gender_ratio_excl_2024)


```

# ITI Berhampur 
```{r}
df_ITI_Berhampur = df_ITI %>% filter(ReportedInstitute == "ITI Berhampur, Ganjam")

# Student Composition (M vs F) over time
gender_ITI_Berhampur = df_ITI_Berhampur %>% 
  group_by(Year,Gender) %>%
  summarise(Count = n()) %>%
  spread(key = Gender,value = Count, fill = 0) %>%
  mutate(Total = Male + Female, PercentFemale = (Female / Total) * 100)
view(gender_ITI_Berhampur)
# Trade composition - Most popular and unpopular among female 

# 


```


# Replicating DTET & SCTEVT Dashboard

```{r}
# install and load packages 
install.packages("pivottabler")
install.packages("openxlsx") # For exporting to Excel

library(pivottabler)
library(openxlsx)


## Intake by ITI - Total
pt <- PivotTable$new()
pt$addData(df_ITI)
pt$addColumnDataGroups("Gender")  # Column by 
pt$addRowDataGroups("Year",addTotal=FALSE)    # Row by 
# Define calculation for the total count (absolute count)
pt$defineCalculation(calculationName = "TotalStudents", 
                     caption = "Student Count", 
                     summariseExpression = "n()")


# Render the pivot table
pt$renderPivot()
# Convert it to a df
iti_students <- as.data.frame(pt$asDataFrame())

## Intake by ITI - Gov
pt <- PivotTable$new()
pt$addData(df_ITI_gov)
pt$addColumnDataGroups("Gender")  # Column by 
pt$addRowDataGroups("Year",addTotal=FALSE)    # Row by 
# Define calculation for the total count (absolute count)
pt$defineCalculation(calculationName = "TotalStudents", 
                     caption = "Student Count", 
                     summariseExpression = "n()")


# Render the pivot table
pt$renderPivot()
# Convert it to a df
iti_students_gov <- as.data.frame(pt$asDataFrame())

## Intake by ITI - Pvt
pt <- PivotTable$new()
pt$addData(df_ITI_pvt)
pt$addColumnDataGroups("Gender")  # Column by 
pt$addRowDataGroups("Year",addTotal=FALSE)    # Row by 
# Define calculation for the total count (absolute count)
pt$defineCalculation(calculationName = "TotalStudents", 
                     caption = "Student Count", 
                     summariseExpression = "n()")


# Render the pivot table
pt$renderPivot()
# Convert it to a df
iti_students_pvt <- as.data.frame(pt$asDataFrame())


## Intake by Diploma - Total
pt <- PivotTable$new()
pt$addData(df_Diploma)
pt$addColumnDataGroups("Gender")  # Column by 
pt$addRowDataGroups("Year",addTotal=FALSE)    # Row by 
# Define calculation for the total count (absolute count)
pt$defineCalculation(calculationName = "TotalStudents", 
                     caption = "Student Count", 
                     summariseExpression = "n()")


# Render the pivot table
pt$renderPivot()
# Convert it to a df
diploma_students <- as.data.frame(pt$asDataFrame(includeHeaderValues=TRUE))

## Intake by Diploma - Gov
df_Diploma_gov = df_Diploma %>% filter(TypeofInstitute == "Govt.")
pt <- PivotTable$new()
pt$addData(df_Diploma_gov)
pt$addColumnDataGroups("Gender")  # Column by 
pt$addRowDataGroups("Year",addTotal=FALSE)    # Row by 
# Define calculation for the total count (absolute count)
pt$defineCalculation(calculationName = "TotalStudents", 
                     caption = "Student Count", 
                     summariseExpression = "n()")


# Render the pivot table
pt$renderPivot()
# Convert it to a df
diploma_students_gov <- as.data.frame(pt$asDataFrame())

## Intake by Diploma - Pvt
df_Diploma_pvt = df_Diploma %>% filter(TypeofInstitute == "Pvt.")
pt <- PivotTable$new()
pt$addData(df_Diploma_pvt)
pt$addColumnDataGroups("Gender")  # Column by 
pt$addRowDataGroups("Year",addTotal=FALSE)    # Row by 
# Define calculation for the total count (absolute count)
pt$defineCalculation(calculationName = "TotalStudents", 
                     caption = "Student Count", 
                     summariseExpression = "n()")


# Render the pivot table
pt$renderPivot()
# Convert it to a df
year_data <- as.character(unlist(pt$getDistinctRowGroups()))


diploma_students_pvt <- pt$asDataFrame()

# Combine row headers with pivot data
diploma_students_pvt <- cbind(Year=year_data, diploma_students_pvt)


# Export Student count to excel file
student_wb = createWorkbook()
addWorksheet(student_wb, "ITI - Total")
writeData(student_wb, "ITI - Total", iti_students)

addWorksheet(student_wb, "ITI - Gov")
writeData(student_wb, "ITI - Gov", iti_students_gov)

addWorksheet(student_wb, "ITI - Pvt")
writeData(student_wb, "ITI - Pvt", iti_students_pvt)

addWorksheet(student_wb, "Polytechnic - Total")
writeData(student_wb, "Polytechnic - Total", diploma_students)

addWorksheet(student_wb, "Polytechnic - Gov")
writeData(student_wb, "Polytechnic - Gov", diploma_students_gov)

addWorksheet(student_wb, "Polytechnic - Pvt")
writeData(student_wb, "Polytechnic - Pvt", diploma_students_pvt)

# Save the workbook to a file
saveWorkbook(student_wb, file = "C:\\Users\\Admin\\Documents\\GitHub\\sams\\output\\Student Count.xlsx", overwrite = TRUE)



student_count_ITI = df_ITI %>%
  group_by(Year) %>%
  summarise(
    Student_count = n()) %>%
  arrange(Year)
view(student_count_ITI)

student_count_PDIS = df_PDIS %>%
  group_by(Year) %>%
  summarise(
    Student_count = n()) %>%
  arrange(Year)
view(student_count_PDIS)




# Intake by District


# Intake by Type 


# Intake by Branch

# Institute Count 
count_ITI = df_ITI %>%
  group_by(Year) %>%
  summarise(
    total_institutes = n_distinct(ReportedInstitute),
    gov_institutes = n_distinct(ReportedInstitute[TypeofInstitute == "Govt."]),
    pvt_institutes = n_distinct(ReportedInstitute[TypeofInstitute == "Pvt."])
  ) %>%
  arrange(Year)
view(count_ITI)

count_Poly = df_Diploma %>%
  group_by(Year) %>%
  summarise(
    total_institutes = n_distinct(ReportedInstitute),
    gov_institutes = n_distinct(ReportedInstitute[TypeofInstitute == "Govt."]),
    pvt_institutes = n_distinct(ReportedInstitute[TypeofInstitute == "Pvt."])
  ) %>%
  arrange(Year)
view(count_Poly)

count_PDIS = df_PDIS %>%
  group_by(Year) %>%
  summarise(
    total_institutes = n_distinct(ReportedInstitute),
    gov_institutes = n_distinct(ReportedInstitute[TypeofInstitute == "Govt."]),
    pvt_institutes = n_distinct(ReportedInstitute[TypeofInstitute == "Pvt."])
  ) %>%
  arrange(Year)
view(count_PDIS)


# Export institute count to an excel file
count_wb = createWorkbook()
addWorksheet(count_wb, "ITI")
writeData(count_wb, "ITI", count_ITI)

addWorksheet(count_wb, "Polytechnic")
writeData(count_wb, "Polytechnic", count_Poly)

addWorksheet(count_wb, "PDIS")
writeData(count_wb, "PDIS", count_PDIS)

# Save the workbook to a file
saveWorkbook(count_wb, file = "C:\\Users\\Admin\\Documents\\GitHub\\sams\\output\\Institute Count.xlsx", overwrite = TRUE)


```



# Maps

```{r}

```


# Reference Code for Map
```{r}
# Install necessary packages if not already installed
# install.packages(c("sf", "tidyverse", "tmap"))


library(sf)
library(tmap)

# Step 1: Load the shapefile of Odisha districts
# Replace with the actual path to your shapefile
odisha_districts <- st_read("C:/Users/Admin/Downloads/odisha.kml")

# Step 3: Count the number of ITIs in each district
iti_count <- iti_list_2024_25 %>%
  group_by(District_Name) %>%
  summarise(num_itis = n())

iti_count = iti_count %>% mutate(District_Name = toupper(District_Name))

iti_count <- iti_count %>%
  mutate(District_Name = case_when(
    District_Name == "ANGUL" ~ "ANUGUL",
    District_Name == "BOLANGIR" ~ "BOLANGIR (BALANGIR)",
    District_Name == "BALASORE" ~ "BALASORE (BALESHWAR)",
    District_Name == "JAJPUR" ~ "JAJAPUR",
    District_Name == "KENDRAPARA" ~ "KENDRAPARHA",
    District_Name == "KEONJHAR" ~ "KEONJHAR (KENDUJHAR)",
    District_Name == "KHURDA" ~ "KHORDHA",
    District_Name == "NAWARANGPUR" ~"NABARANGAPUR",
    District_Name == "NUAPADA" ~"NUAPARHA",
    District_Name == "RAYAGADA" ~"RAYAGARHA",
    District_Name == "SONEPUR" ~"SUBARNAPUR",
    District_Name == "BOUDH" ~"BAUDH (BAUDA)",
    TRUE ~ District_Name  # Keep other district names unchanged
  ))


# Step 4: Merge the ITI count data with the Odisha districts shapefile
odisha_districts <- odisha_districts %>%
  left_join(iti_count, by = c("Name" = "District_Name"))

breaks = c(1,10,20,30,40,51)
# Step 5: Visualize the map
# Plotting with tmap
tmap_mode("plot")  # Change to "view" for interactive mode
tm_shape(odisha_districts) +
  tm_polygons("num_itis", 
              title = "Number of ITIs", 
              palette = "Blues", 
              breaks = breaks,
              n = 5) +
  tm_layout(main.title = "Number of ITIs in Each District of Odisha",
            main.title.size = 1.2) +
  tm_borders(lwd = 3, col = "black")

## For NVCT
odisha_iti_seats_all$ITI.Count = as.numeric(odisha_iti_seats_all$ITI.Count)

odisha_iti_seats_all = odisha_iti_seats_all %>% mutate(State.District = toupper(State.District))

odisha_districts <- odisha_districts %>%
  left_join(odisha_iti_seats_all, by = c("Name" = "State.District"))

# change count of ITIs in SUBARNAPUR to 2 and trade.count to 7
odisha_districts[28, "ITI.Count"] <- 2
odisha_districts[28, "Trade.Count"] = 7

odisha_districts = odisha_districts[-29,]
odisha_districts$ITI.Count = as.numeric(odisha_districts$ITI.Count)

odisha_iti_seats_all <- odisha_iti_seats_all %>%
  mutate(State.District = case_when(
    State.District == "BARGARH" ~ "BARAGARH",
    State.District == "BALANGIR" ~ "BOLANGIR (BALANGIR)",
    State.District == "BALASORE" ~ "BALASORE (BALESHWAR)",
    State.District == "JAGATSINGHAPUR" ~ "JAGATSINGHPUR",
    State.District == "KENDRAPARA" ~ "KENDRAPARHA",
    State.District == "KENDUJHAR" ~ "KEONJHAR (KENDUJHAR)",
    State.District == "KHORDHA" ~ "KHORDHA",
    State.District == "NABARANGPUR" ~"NABARANGAPUR",
    State.District == "NUAPADA" ~"NUAPARHA",
    State.District == "RAYAGADA" ~"RAYAGARHA",
    State.District == "BOUDH" ~"BAUDH (BAUDA)",
    TRUE ~ State.District  # Keep other district names unchanged
  ))

# Visualize NCVT figures
tmap_mode("plot")  # Change to "view" for interactive mode
tm_shape(odisha_districts) +
  tm_polygons("ITI.Count", 
              title = "Number of ITIs", 
              palette = "Blues", 
              breaks = breaks,
              n = 5) +
  tm_layout(main.title = "Number of ITIs in Each District of Odisha",
            main.title.size = 1.2) +
  tm_borders(lwd = 3, col = "black")


# Difference between SAMS and NCVT
odisha_districts$diff_trade_count = odisha_districts$ITI.Count - odisha_districts$num_itis

# Visualize the difference
tmap_mode("plot")  # Change to "view" for interactive mode
tm_shape(odisha_districts) +
  tm_polygons("diff_trade_count", 
              title = "Difference in ITIs", 
              palette = "Blues", 
              style = "quantile",
              n = 5) +
  tm_layout(main.title = "Difference in Number of ITIs SAMS and NCVT",
            main.title.size = 1.2) +
  tm_borders(lwd = 3, col = "black")

# breaks for seat utilization
breaks_utilization = c(68.0,80,85.0,90.0,95.0,100.0)

# Visualize Seat utilization of "Previous Year" i.e. 2021
tmap_mode("view")  # Change to "view" for interactive mode
tm_shape(odisha_districts) +
  tm_polygons("Seat.Utilization.Previous", 
              title = "Seat Utilization (%)", 
              palette = "Blues", 
              breaks = breaks_utilization,
              popup.vars = c("Seat Utilization Rate" = "Seat.Utilization.Previous"),
              n = 5) +
  tm_layout(main.title = "Seat Utilization Rate in ITIs",
            main.title.size = 1.2) +
  tm_borders(lwd = 3, col = "black")

```






