---
title: "SAMS Data Analysis Draft"
output: html_document
date: "`r Sys.Date()`"
---

```{r, echo=FALSE}
install.packages("DBI")
install.packages("RSQLite")
install.packages("kableExtra")

```

# Load the libraries

```{r, echo=FALSE}
library(RSQLite)
library(DBI)
library(dplyr)
library(tidyverse)
library(kableExtra)
```


# Application - Enrollment 

```{r, echo=FALSE}
# Check for disrepencies in applied_status and admission and enrollment
table(student_data_iti$admission_status,student_data_iti$enrollment_status)

# Applied_status and Admission_status Matrix
table(student_data_iti$applied_status, student_data_iti$admission_status)
# Above - (Yes,Yes) means applied and accepted. (Yes,No) means applied and rejected. (No,No) means did not apply so rejected. (No,Yes) is a puzzle. How can one not apply and get in? 


```
There is four columns with information on application and enrollment: 
application_status: Accepted/Pending 
applied_status: Yes/No
enrollment_status: Yes/No
admission_status: Yes/No

enrollment_status and admission_status are the same.
application_status and applied_status are equivalent (Accepted = Yes; Pending = No)

No in applied status probably means those who were not able to complete applying or did not pay the fees. (Confirm with CSM once) 

To calculate the count who did not get admission - Yes in applied status; No in enrollment_status
It is not sure among those who did not get admission whether they were rejected or did not accept. 

A categorical dummy can be made to indicate rejected applicants and accepted applicants. 
After that, a comparison between the two groups can be done based on various variables. 


## ITI Student Data from 2017 to 2024
```{r, echo=FALSE}
con <- dbConnect(RSQLite::SQLite(), "C:/Users/Admin/Documents/GitHub/sams/data/raw/sams.db")

# Query data for ITI module for years 2017 to 2023
query <- "SELECT * FROM students WHERE module = 'ITI' AND year BETWEEN 2017 AND 2024;"

# Send query to the database
result <- dbSendQuery(con, query)

# Fetch the relevant data
student_data_iti <- dbFetch(result)

# Clear the result
dbClearResult(result)
dbDisconnect(con)

# Admitted students only 
student_iti_admis = student_data_iti %>%
  filter(admission_status == "Yes")

```


```{r,eval=FALSE}
## Check the admission count each year
table(student_iti_admis$year)

student_iti_admis = student_data_iti %>%
  filter(admission_status == "Yes")
table(student_iti_admis$year)


iti_admis_count = student_iti_admis %>%
  group_by(year) %>%
  summarise(Admitted_students_iti = n())
view(iti_admis_count)


# Create the 'final_status' variable
#student_data_iti <- student_data_iti %>%
#  mutate(final_status = case_when(
#    admissionstatus == "Yes" ~ "accepted",
#    admissionstatus == "No" & appliedstatus == "Yes" ~ "rejected",
#    admissionstatus == "No" & appliedstatus == "No" ~ "not applied"
#  ))


```


## Diploma Student Data from 2018 to 2024 
```{r, include=FALSE,echo=FALSE}
con <- dbConnect(RSQLite::SQLite(), "C:/Users/Admin/Documents/GitHub/sams/data/raw/sams.db")

# Query data for ITI module for years 2017 to 2023
query <- "SELECT * FROM students WHERE module = 'Diploma' AND year BETWEEN 2018 AND 2024;"

# Send query to the database
result <- dbSendQuery(con, query)

# Fetch the relevant data
student_data_dip <- dbFetch(result)

# Clear the result
dbClearResult(result)
dbDisconnect(con)



## Check the admission count each year
student_dip_admis = student_data_dip[!is.na(student_data_dip$sams_code),]


dip_admis_count = student_dip_admis %>%
  group_by(year) %>%
  summarise(Admitted_students_Diploma = n())

```

```{r,eval=FALSE}
table(student_data_dip$admission_status,student_data_dip$applied_status)

table(student_dip_admis$year)

student_dip_admis = student_data_dip %>%
  filter(admission_status == "Yes")
table(student_dip_admis$year)

table(student_data_dip$highest_qualification)


```


## Check for duplicates in admission data for ITI and Diploma
```{r, echo=FALSE}
## ITI
iti_duplicates <- student_iti_admis %>%
  group_by(aadhar_no,year) %>%
  summarize(num_duplicates = n()) %>%
  arrange(num_duplicates)
#head(iti_duplicates)

duplicate_barcodes <- iti_duplicates %>%
 filter(num_duplicates > 1) %>%   # Filter rows where num_duplicates are greater than 1
 select(aadhar_no)           # Select only those Barcodes


barcode_dupli_list = student_iti_admis %>% filter(aadhar_no %in% duplicate_barcodes$aadhar_no)
#head(barcode_dupli_list)


## Diploma
dip_duplicates <- student_dip_admis %>%
  group_by(aadhar_no,year) %>%
  summarize(num_duplicates = n()) %>%
  arrange(num_duplicates)
#head(dip_duplicates)

duplicate_barcodes <- dip_duplicates %>%
 filter(num_duplicates > 1) %>%   # Filter rows where num_duplicates are greater than 1
 select(aadhar_no)           # Select only those Barcodes


barcode_dupli_list = student_dip_admis %>% filter(aadhar_no %in% duplicate_barcodes$aadhar_no)
#head(barcode_dupli_list)


```

```{r,eval=FALSE}
student_data_iti_excl_2017 = student_data_iti %>%
  filter(year != "2017")

# Filter out those who have a higher than 10th Highest Qualification
student_data_iti_tenth_excl_2017 = student_data_iti_excl_2017 %>%
  filter(highest_qualification == "10th")
# Couldn't do for Diploma cause "highest_qualification" is empty (NA)


# Create a function to calculate the counts for each category
process_data <- function(data) {
  data %>%
    group_by(year) %>%
    summarise(
      applied = sum(applied_status == "Yes", na.rm = TRUE),
      admitted = sum(admission_status == "Yes", na.rm = TRUE),
      not_admitted = sum(admission_status == "No" & applied_status == "Yes", na.rm = TRUE)
    )
}


# Process ITI and Diploma data separately
iti_summary <- process_data(student_data_iti_tenth_excl_2017)
diploma_summary <- process_data(student_data_dip)

# Filter for the year 2023
iti_2023 <- iti_summary %>% filter(year == 2023)
diploma_2023 <- diploma_summary %>% filter(year == 2023)

```


## Pipeline Table of 10th - +2 - ITI/Diploma
```{r, echo=FALSE}
# Exclude 2017 ITI

# Import 10th figures
tenth_count = read_excel("C:/Users/Admin/Documents/GitHub/sams/data/raw/tenth_plustwo_odisha.xlsx")

# Summarize the data by year, calculating the total pass, total fail, and total batch count
total_tenth_count <- tenth_count %>%
  group_by(Year, Gender) %>%
  summarise(
    total_pass = sum(Pass, na.rm = TRUE),
    total_fail = sum(Fail, na.rm = TRUE),
    total_batch = total_pass + total_fail
  ) %>%
  ungroup() %>%
  group_by(Year) %>%
  summarise(
    total_pass = sum(total_pass),
    total_fail = sum(total_fail),
    total_batch = total_pass + total_fail
  )
view(total_tenth_count)

# Tenth passing batch only 2023
total_tenth_count_2023 = total_tenth_count %>% filter(Year == 2023)

# Import admission count for +2 in from 2011 to 2023 from SAMS Dashboard
plus_two_admis <- read_excel("C:/Users/Admin/Documents/GitHub/sams/data/raw/tenth_plustwo_odisha.xlsx", 
    sheet = "plus_two")

plus_two_admis <- plus_two_admis %>% rename(year = Year)
total_tenth_count <- total_tenth_count %>% rename(year = Year)

total_tenth_count <- total_tenth_count %>% rename(total_tenth_batch = total_batch)
plus_two_admis <- plus_two_admis %>% rename(total_plus_two_admitted = student_admitted)

view(total_tenth_count)
# Select only the columns you need from each dataframe
df1_selected <- plus_two_admis %>% select(year, total_plus_two_admitted)
df2_selected <- total_tenth_count %>% select(year, total_tenth_batch)

# Perform full_join based on 'Year'
pipeline_table <- df2_selected %>%
  full_join(df1_selected, by = "year") %>%
  full_join(iti_admis_count, by = "year") %>%
  full_join(dip_admis_count, by = "year") %>%
  arrange(year) %>%
  mutate(total_plus_two_admitted = gsub(",", "", total_plus_two_admitted)) %>%
  mutate(total_plus_two_admitted = as.numeric(total_plus_two_admitted))



```


## Sankey Plot of 10th - ITI/Diploma
```{r, eval=FALSE, echo=FALSE}
# Prepare df for Sankey
sankey_data <- data.frame(
  source = c("10th Batch", "10th Batch", "10th Batch","10th Batch", "Applied ITI", "Applied ITI", "Applied Diploma", "Applied Diploma"),
  target = c("Admitted in +2", "Applied ITI", "Applied Diploma", "Others", "Applied and Admitted ITI", "Applied and Not Admitted ITI", "Applied and Admitted Diploma", "Applied and Not Admitted Diploma"),
  value = c(
    Plus_two_admis_2023, # Admitted in +2
    iti_2023$applied,   # Applied ITI
    diploma_2023$applied,   # Applied Diploma
    total_tenth_count_2023$total_batch - iti_2023$applied - diploma_2023$applied - Plus_two_admis_2023,  # Others
    iti_2023$admitted,  # Applied and Admitted ITI
    iti_2023$not_admitted,  # Applied and Not Admitted ITI
    diploma_2023$admitted,  # Applied and Admitted Diploma
    diploma_2023$not_admitted  # Applied and Not Admitted Diploma
  )
)
view(sankey_data)

# Plot Sankey
library(networkD3)

# Create Nodes dataframe
nodes <- data.frame(name = unique(c(sankey_data$source, sankey_data$target)))

# Convert source and target to indices
sankey_data$source <- match(sankey_data$source, nodes$name) - 1
sankey_data$target <- match(sankey_data$target, nodes$name) - 1


# Create Sankey plot
sankey <- sankeyNetwork(Links = sankey_data, Nodes = data.frame(name = unique(c(sankey_data$source, sankey_data$target))),
                        Source = "source", Target = "target", Value = "value", NodeID = "name", fontSize = 12, nodeWidth = 30)

str(sankey_data)

# Plot the Sankey diagram
sankey

```


## Migration
```{r, echo=FALSE}
# Diagnose 
#table(student_iti_admis$district,student_iti_admis$institute_district)
#table(student_dip_admis$district,student_dip_admis$institute_district)

#n_distinct(student_dip_admis$district)
#n_distinct(student_data_iti$institute_district)
# Number of districts in "institute_district" for ITIs is 30 (The number of districts in Odisha)


## ITI Students 
# Note - Only migration within Odisha state
# Count the students who leave their home district
ITI_leave_dist = student_iti_admis %>%
  filter(!district %in% c("Other", "Skill Department (DTET)", "Odisha")) %>% # Exclude these
  group_by(district,year,institute_district,gender,typeof_institute) %>%
  summarise(leave_home_dist = sum(district != institute_district)) %>%
  ungroup() 

# How many leave their home district to study in 2023


## Diploma Students 
# Note - Only migration within Odisha state
# Count the students who leave their home district
Dip_leave_dist = student_dip_admis %>%
  filter(district != "Other") %>% # Exclude these
  group_by(district,year,institute_district,gender,typeof_institute) %>%
  summarise(leave_home_dist = sum(district != institute_district)) %>%
  ungroup() 

## Aggregate - How many leave their home district to study 
# ITI
ITI_total_leave_dist = student_iti_admis %>%
  filter(!district %in% c("Other", "Skill Department (DTET)", "Odisha")) %>% # Exclude these
  group_by(year) %>%
  summarise(total_leaving_home_dist = sum(district != institute_district)) %>%
  ungroup() 

print(ITI_total_leave_dist)
# Diploma
Dip_total_leave_dist = student_dip_admis %>%
  filter(district != "Other") %>% # Exclude these
  group_by(year) %>%
  summarise(total_leaving_home_dist = sum(district != institute_district)) %>%
  ungroup() 

## Aggregate by Gender - How many leave their home district to study 
# ITI
ITI_gender_leave_dist = student_iti_admis %>%
  filter(!district %in% c("Other", "Skill Department (DTET)", "Odisha")) %>% # Exclude these
  group_by(year, gender) %>%
  summarise(total_leaving_home_dist = sum(district != institute_district)) %>%
  ungroup() 
# Pivot the data frame
pivoted_ITI_gender_leave_dist <- ITI_gender_leave_dist %>%
  pivot_wider(names_from = gender, values_from = total_leaving_home_dist, values_fill = 0)
# Remove Transgender
pivoted_ITI_gender_leave_dist <- pivoted_ITI_gender_leave_dist %>%
  select(-Transgender)
# Calculate Female Share
pivoted_ITI_gender_leave_dist <- pivoted_ITI_gender_leave_dist %>%
  mutate(Female_Share_Percent = Female / (Male + Female) * 100)
# Calculate Total Students
pivoted_ITI_gender_leave_dist <- pivoted_ITI_gender_leave_dist %>%
  mutate(Total_Students = Male + Female) %>%
  select(year, Total_Students, Male, Female, Female_Share_Percent)  # Reorder columns


# Diploma
Dip_gender_leave_dist = student_dip_admis %>%
  filter(district != "Other") %>% # Exclude these
  group_by(year, gender) %>%
  summarise(total_leaving_home_dist = sum(district != institute_district)) %>%
  ungroup() 
pivoted_Dip_gender_leave_dist <- Dip_gender_leave_dist %>%
  pivot_wider(names_from = gender, values_from = total_leaving_home_dist, values_fill = 0)
# Remove Transgender
pivoted_Dip_gender_leave_dist <- pivoted_Dip_gender_leave_dist %>%
  select(-Transgender)
# Calculate Female Share
pivoted_Dip_gender_leave_dist <- pivoted_Dip_gender_leave_dist %>%
  mutate(Female_Share_Percent = Female / (Male + Female) * 100)
# Calculate Total Students
pivoted_Dip_gender_leave_dist <- pivoted_Dip_gender_leave_dist %>%
  mutate(Total_Students = Male + Female) %>%
  select(year, Total_Students, Male, Female, Female_Share_Percent)  # Reorder columns




## Aggregate by Home District - How many leave their home district to study 
# ITI
ITI_home_leave_dist = student_iti_admis %>%
  filter(!district %in% c("Other", "Skill Department (DTET)", "Odisha")) %>% # Exclude these
  group_by(year, district) %>%
  summarise(total_leaving_home_dist = sum(district != institute_district)) %>%
  ungroup() 

# Diploma
Dip_home_leave_dist = student_dip_admis %>%
  filter(district != "Other") %>% # Exclude these
  group_by(year, district) %>%
  summarise(total_leaving_home_dist = sum(district != institute_district)) %>%
  ungroup() 

## Aggregate by Institute District - How many leave their home district to study 
# ITI
ITI_insti_leave_dist = student_iti_admis %>%
  filter(!district %in% c("Other", "Skill Department (DTET)", "Odisha")) %>% # Exclude these
  group_by(year, institute_district) %>%
  summarise(total_leaving_home_dist = sum(district != institute_district)) %>%
  ungroup() 

# Diploma
Dip_insti_leave_dist = student_dip_admis %>%
  filter(district != "Other") %>% # Exclude these
  group_by(year, institute_district) %>%
  summarise(total_leaving_home_dist = sum(district != institute_district)) %>%
  ungroup() 


```


```{r, results='asis'}
pipeline_table %>%
  filter(year >= 2017) %>%
  knitr::kable(caption = "Table 1: Vocational Training Pipeline in Odisha", digits = 1) %>%
  kableExtra::kable_styling(bootstrap_options = "striped",full_width = FALSE)

pivoted_ITI_gender_leave_dist %>%
  knitr::kable(caption = "Table 2: ITI Students Leaving Home District", digits = 1) %>%
  kableExtra::kable_styling(bootstrap_options = "striped",full_width = FALSE)

pivoted_Dip_gender_leave_dist %>%
  knitr::kable(caption = "Table 3: Diploma Students Leaving Home District", digits = 1) %>%
  kableExtra::kable_styling(bootstrap_options = "striped",full_width = FALSE)

```



```{r}
# Check if the columns are identical 
are_identical <- identical(student_iti_admis$year, student_iti_admis$academic_year)

# Print result
print(are_identical)  


```


