---
title: "SAMS Data Analysis"
author: "Data, Policy, and Innovation Centre"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
always_allow_html: true
output:
  word_document:
    toc: true
    toc_depth: 3
    fig_caption: true
    keep_md: false           # Choose whether to keep the intermediate .md file
    df_print: paged         # Prints data frames as paginated tables
    reference_docx: "template.docx" 
---

```{r Load configuration, include=FALSE, eval=FALSE, echo=FALSE}
library(here)
 rm(list = ls())

here::i_am("README.md")
here()
source(file.path(here(),"scripts","config.R"))


```

```{r Libraries, include=FALSE, echo=FALSE}
library(tidyverse)
library(arrow)
library(ggplot2)
library(openxlsx)
library(kableExtra)
library(knitr)
library(stargazer)
library(RColorBrewer)
```

```{r Load processed data, include=FALSE, echo=FALSE}
# Processed data
iti_marks_and_cutoffs <- read_parquet(get_path("iti_marks_and_cutoffs"))
iti_vacancies <- read_parquet(get_path("iti_vacancies"))
iti_marks_and_cutoffs_2023 <- iti_marks_and_cutoffs %>% filter(academic_year == 2023)
```


```{r Load interim data, include=FALSE, echo=FALSE}
# Interim data
iti_enrollments <- read_parquet(get_path("iti_enrollments"))
iti_enrollments_2023 <- iti_enrollments %>% filter(academic_year == 2023)
iti_cutoffs <- read_parquet(get_path("iti_institutes_cutoffs"))

```

```{r Add type of institutes data to cutoffs, include=FALSE, echo=FALSE}
iti_cutoffs <- iti_enrollments %>% distinct(sams_code,type_of_institute) %>% 
  right_join(iti_cutoffs, by="sams_code")
iti_cutoffs_2023 <- iti_cutoffs %>% filter(academic_year == 2023)


```

# Executive summary

*Fill in later*

\newpage

# An overview of SAMS

The Student Academic Management System (SAMS) is a digital platform developed and deployed by the Government of Odisha to streamline the admission and administration processes across various levels of education in the state. It serves as a one-stop solution for managing academic workflows, primarily focusing on admissions, academic tracking, and data management. Initially introduced in 2010 for higher secondary education, SAMS has expanded to encompass higher education and vocational training, making it a pivotal tool for educational governance in Odisha.

## Key features

* Centralized Admission Process
      SAMS facilitates online applications for admissions across higher secondary schools, colleges, ITIs, and vocational training institutions. This unified system ensures:

    * Transparency in the selection process.
    * Equal access to educational opportunities.
    * Minimization of manual intervention and associated errors.

* Multi-Tier Educational Coverage
    The platform covers the following educational sectors:

    * Higher Secondary Education (Class XI-XII): Admissions into schools offering Science, Commerce, and Arts streams.
    * Higher Education (Undergraduate Programs): Colleges offering general and professional degree programs.
    * Vocational and Skill-Based Training: Industrial Training Institutes (ITIs) and Diploma courses in engineering and other trades.
    * Post-Diploma and Advanced Courses: Specialized technical and professional courses.

* End-to-End Digital Process
  SAMS integrates and digitizes the student application experience:

    * Digital registration and application submission.
    * Real-time merit-based seat allocation.

## The SAMS database as a resource for policy analysis

SAMS has proven to be a transformative tool for streamlining admissions across various levels of education in Odisha. It has successfully automated processes, enhanced transparency, and lessened the administrative burden on individual institutions. However, while the platform hosts a rich repository of data on student demographics, academic performance, enrollment patterns, and institutional capacities, this information remains underutilized for analytics-driven decision-making and policy formulation.

The primary focus of SAMS has been operational, such as managing admissions and tracking seats across institutions. However, the data generated from these processes is rarely used to identify trends or address systemic issues in education. Despite the capability to generate real-time reports, the data is seldom used to monitor emerging trends or respond proactively to challenges, such as the underutilization of seats in certain programs and regional / socioeconomic disparities in access to education. 

# The state of vocational training in Odisha 

*Add introduction describing vocational training in Odisha and describe how the SAMS analysis showcases the richness of the SAMS data and highlights its ability to help policymakers and adminstrators make data-driven decisions*

##  Overview

*Describe the variables captured in the data*

### Enrollment

```{r Load basic enrollment tables, echo=FALSE, include=FALSE}

path <- get_exhibit("students_enrollment_basics")
sheet_names <- getSheetNames(path)
enrollments_over_time <- read.xlsx(
  path,
  sheet_names[1]
)
iti_time_type <- read.xlsx(
  path,
  sheet_names[2]
)
diploma_time_type <- read.xlsx(
  path,
  sheet_names[3]
)
```



* Exhibit: Table/Figure on enrollment over time in ITIs and Diploma institutes

```{r Enrollment over time, echo=FALSE, include=FALSE}
html_table <- enrollments_over_time %>% kable( caption = "Enrollment over time") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)

# Check why this table doesn't match tables below

```


* Exhibit: Table/Figure on enrollment over time (%) by pvt vs govt for ITIs

```{r Enrollment over time in ITIs by type, echo=FALSE}
html_table <- iti_time_type %>% kable( caption = "Enrollment at ITIs over time by type of institute") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)

```


* Exhibit: Table/Figure on enrollment over time (%) by pvt vs govt for Diplomas


```{r Enrollment over time in Diplomas by type, echo=FALSE}
html_table <- diploma_time_type %>% kable( caption = "Enrollment in Diplomas over time by type of institute") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)

```



### Institutes

```{r Load institutes basics tables, include=FALSE, echo=FALSE}
path <- get_exhibit("institutes_basics")
sheet_names <- getSheetNames(path)
num_iti_over_time <- read.xlsx(
  path,
  sheet_names[1]
)
num_diploma_over_time <- read.xlsx(
  path,
  sheet_names[2]
)
all_over_time <- read.xlsx(
  path,
  sheet_names[3]
)

num_itis_time_type <- read.xlsx(
  path,
  sheet_names[4]
)

num_diploma_time_type <- read.xlsx(
  path,
  sheet_names[5]
)

top_10_iti_enrollment <- read.xlsx(
  path,
  sheet_names[6]
)

top_10_diploma_enrollment <- read.xlsx(
  path,
  sheet_names[7]
)


```


* Exhibit: Table on institutes over time (ITI/Diploma)



```{r Num ITIs over time, echo=FALSE}
html_table <- num_iti_over_time %>% kable( caption = "Num. ITIs over time") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)

```


```{r Num Diploma over time, echo=FALSE}
html_table <- num_diploma_over_time %>% kable( caption = "Num. Diploma institutes over time") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)

```



* Exhibit: Table on institutes over time by pvt vs govt for ITIs


```{r ITI enrollment over time by type, echo=FALSE}
html_table <- num_itis_time_type %>% kable( caption = "Num. ITIs over time by type of institute") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)

```


* Exhibit: Table on institutes over time by pvt vs govt for Diplomas



```{r Diploma enrollment over time by type, echo=FALSE}
html_table <- num_diploma_time_type %>% kable( caption = "Num. Diploma institutes over time by type of institute") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)

```



* Exhibit: Top 10 institutes by enrollment in 2023 for ITIs

```{r Top 10 ITIs in 2023 by enrollment, echo=FALSE}
html_table <- top_10_iti_enrollment %>% kable( caption = "Top 10 ITIs by enrollment in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)

```


* Exhibit: Top 10 institutes by enrollment in 2023 for Diplomas


```{r Top 10 Diplomas in 2023 by enrollment, echo=FALSE}
html_table <- top_10_diploma_enrollment %>% kable( caption = "Top 10 ITIs by enrollment in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)

```




* Exhibit: Summary stats on enrollment in institutions over time by pvt vs govt


### Branches and trades

```{r Load branches and trades tables, echo=FALSE, include=FALSE}
path <- get_exhibit("trades_and_branches_basics")
sheet_names <- getSheetNames(path)
trades_over_time <- read.xlsx(
  path,
  sheet_names[1]
)
branches_over_time <- read.xlsx(
  path,
  sheet_names[2]
)
top_10_itis_num_trades <- read.xlsx(
  path,
  sheet_names[3]
)
top_10_diploma_num_branches <- read.xlsx(
  path,
  sheet_names[4]
)
top_10_trades  <- read.xlsx(
  path,
  sheet_names[5]
)

top_10_branches <-  read.xlsx(
  path,
  sheet_names[6]
)

```


* Exhibit: Total number of trades offered by ITIs over time

```{r Trades over time, include=TRUE, echo=FALSE}

html_table <- trades_over_time %>% kable( caption = "Num. ITI trades offered over time") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)


```


* Exhibit: Total number of branches in Diploma institutes over time


```{r Branches over time, include=TRUE, echo=FALSE}

html_table <- branches_over_time %>% kable( caption = "Num. Diploma branches offered over time") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)


```


* Exhibit: The top 10 trades by enrollment at ITIs in 2023

```{r Top 10 trades by enrollment in 2023, include=TRUE, echo=FALSE}

html_table <- top_10_trades %>% kable( caption = "Top 10 ITI trades by enrollment in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)


```


* Exhibit: The top 10 branches by enrollment at Diplomas in 2023


```{r Top 10 branches by enrollment in 2023, include=TRUE, echo=FALSE}

html_table <- top_10_branches %>% kable( caption = "Top 10 Diploma branches by enrollment in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)


```


* Exhibit: The top 10 institutes by num. trades offered in 2023 for ITIs 


```{r Top 10 ITIs by num. trades offered, include=TRUE, echo=FALSE}

html_table <- top_10_itis_num_trades %>% kable( caption = "Top 10 ITIs by num. trades offered in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)


```


* Exhibit: The top 10 institutes by num. branches in 2023 for Diplomas


```{r Top 10 Diplomas by num. branches, include=TRUE, echo=FALSE}

html_table <- top_10_diploma_num_branches %>% kable( caption = "Top 10 Diplomas by number of branches in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)


```


## The vocational training pipeline

```{r Load pipeline data, echo=FALSE, include=FALSE}
path <- 

```


* Exhibit: Student enrollment after class 10 over time

* Exhibit: Gap between 10th graduation and enrollment

## Student demographics

* Exhibit: Share of students by gender over time for ITIs

* Exhibit: Share of students by gender over time for Diplomas

* Exhibit: Students by social category over time for ITIs

* Exhibit: Students by social category over time for Diplomas

* Exhibit: Crosstabulation of students by gender and social category in 2023 for ITIs

* Exhibit: Crosstabulation of students by gender and social category in 2023 for Diploma

* Exhibit: Crosstabulation of students by income and social category in 2023 for ITI

* Exhibit: Crosstabulation of students by income and social category in 2023 for Diploma

* Exhibit: Top 10 trades with highest representation of general category students in 2023

* Exhibit: Top 10 trades with highest representation of SC students in 2023

* Exhibit: Top 10 trades with highest representation of ST students in 2023

* Exhibit: Top 10 trades with highest representation of women and girls in 2023

## Student qualifications and marks

* Exhibit: Number / share of students by highest qualification in ITIs in 2023

* Exhibit: Number / share of students by highest qualification in Diplomas in 2023

* Exhibit: Highest qualification by social category in 2023

* Exhibit: Highest qualification by gender in 2023

* Exhibit: Distribution of 10th grade marks for ITI students

* Exhibits: *Select others from annex*


## Seat utilization and vacancies

* Exhibits: *Select from annex*

## The geography of vocational training

* Exhibits: *Select distance exhibits from annex*

\newpage








<!-- # Goals for the report
The long-term goal of this analysis is to uncover motivations and patterns behind students’ enrollment in vocational training program and uncover factors influencing choice based on gender and social demographics. Of principal interest to us is

* What types of students decide to enroll in vocational training programs?
    * This ideally requires data on enrollment in higher education / 12th grade programs as well

* Are there "overqualified" students enrolling in vocational programs?

* Does distance from training institutes act as a binding constraint on students?
    * Are some students unable to enroll in the trades / institutions they would like because of a long commute?
    * Hostel facilities are likely to play a role in this.

* Are there certain trades that are "oversubscribed", in that the government has not provided enough seats in those trades despite demand?
    * Conversely, are some other trades "undersubscribed", and perhaps resources are better allocated in other trades?

The following exhibits serve as a scoping exercise for the report due at the end of the month. -->

# Annex: Exhibit repository

## Trades and qualifications

```{r Load output tables, include=FALSE, echo=FALSE}
quals_excel_path <- file.path(paths$TABLES_DIR,"trades_by_qualification.xlsx")
sheet_names <- getSheetNames(quals_excel_path)
trades_by_quals <- read.xlsx(
  quals_excel_path,
  sheet_names[1]
)
trades_by_quals_pct <- read.xlsx(
  quals_excel_path,
  sheet_names[2]
) %>% rename(
  `Graduate and above` = Graduate.and.above,
  Trade = reported_branch_or_trade
)
trades_by_gender_pct <- read.xlsx(
  quals_excel_path,
  sheet_names[3]
) %>% rename(
  Trade = reported_branch_or_trade
)

qual_by_gender_pct <- read.xlsx(
   quals_excel_path,
   sheet_names[4]
) %>% rename(
  `Graduate and above` = Graduate.and.above
)

trades_by_soc_pct <- read.xlsx(
  quals_excel_path,
  sheet_names[5]
) %>% rename(
  Trade = reported_branch_or_trade
)

qual_by_soc_pct <- read.xlsx(
  quals_excel_path,
  sheet_names[6]
) %>% 
  rename(
    `Graduate and above` = Graduate.and.above
  )

```


* We are likely to find some trades where students are "overqualified". 

```{r Highest qualifications raw, include=FALSE, echo=FALSE}

html_table <- trades_by_quals %>% kable( caption = "Trades by highest qualifications") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)
 
```

```{r Highest qualifications shares, include=TRUE, echo=FALSE}
html_table <- trades_by_quals_pct %>% head(10) %>% kable( caption = "Top 10 trades with highest share of college graduates") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)

```
* Some trades show a particularly high share of graduates
    * These appear to be female dominated trades
    
```{r Trades by gender, include=TRUE, echo=FALSE}
top_10_trades_by_quals_grad <- trades_by_quals_pct %>% head(10)
html_table <- trades_by_gender_pct %>% right_join(top_10_trades_by_quals_grad,by="Trade") %>% select(Trade,Male,Female, Total.x) %>% rename(Total = Total.x) %>% kable( caption = "Top 10 trades by highest share of college graduates, sorted by gender (%)") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)


```

```{r Qualification by gender, echo=FALSE, include=TRUE}
html_table <- qual_by_gender_pct %>%  kable( caption = "Highest qualification by gender (%)") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)

```


```{r Trades by social category, echo=FALSE, include=TRUE}
html_table <- trades_by_soc_pct %>% head(20) %>%  kable( caption = "Trades by social category in 2023 (%)") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)

```
 * Need to check if heavy sorting into trades is due to the geography of trade offerings.
 
```{r Qualifications by social category, echo=FALSE, include=TRUE}
html_table <- qual_by_soc_pct %>% kable( caption = "Qualifications by social category in 2023 (%)") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)


```
 

## Student marks and institute cutoffs

### Marks

```{r Load tables for marks analysis, echo=FALSE,include=FALSE}
marks_excel_path <- file.path(paths$TABLES_DIR, "marks_cutoffs.xlsx")
sheet_names <- getSheetNames(marks_excel_path)
marks_by_trade <- read.xlsx(
  marks_excel_path,
  sheet_names[1]
  
)
marks_by_demographics <- read.xlsx(
  marks_excel_path,
  sheet_names[2]
)
boards_marks <- read.xlsx(
  marks_excel_path,
  sheet_names[3]
)

cutoffs_trade_demographics <- read.xlsx(
  marks_excel_path,
  sheet_names[4]
)

cutoffs_trade_gender <- read.xlsx(
  marks_excel_path,
  sheet_names[5]
)

```

```{r Distribution of 10th marks, echo=FALSE, include=TRUE, warning=FALSE}

ggplot(iti_marks_and_cutoffs_2023, aes(x = percentage)) +
  geom_histogram(
    binwidth = 1,                 # Set bin width
    fill = "#4C72B0",             # Color for the bars
    color = "black",              # Outline color
    alpha = 0.8                   # Transparency
  ) +
  labs(
    title = "Overall distribution of marks in 2023",
    x = "Marks (%)",
    y = "Frequency"
  ) + 
  geom_vline(
    xintercept = 33,              # Set the x position of the line at 33 marks
    color = "red",                # Color of the line
    linetype = "dashed",          # Line type (dashed)
    size = 1                      # Thickness of the line
  ) + 
  annotate(
    "text", 
    x = 33, y = 4500,                # Position the text below the line
    label = "Passing Percentage: 33%", 
    color = "red",                # Color of the text
    size = 3,                     # Size of the text
    hjust = -0.05,
    vjust = -1
  
  ) + 
  theme_classic() 
  

```

```{r Average marks by trade, echo=FALSE, include=TRUE}
html_table <- marks_by_trade %>% head(20) %>% kable( caption = "Summary statistics on marks by trade") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)



```

```{r Average marks by demographics, echo=FALSE,include=TRUE}
html_table <- marks_by_demographics %>% kable(caption = "Average marks by demographics") %>% 
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)


```



```{r Average marks by gender and category within trade, echo=FALSE, include=FALSE}

iti_marks_gender_cat_reg_data <- iti_marks_and_cutoffs_2023
iti_marks_gender_cat_reg_data$gender <- relevel(factor(iti_marks_gender_cat_reg_data$gender),ref="Male")

iti_marks_gender_cat_reg_data$social_category <- relevel(factor(iti_marks_gender_cat_reg_data$social_category),ref="UR")

iti_marks_gender_cat_reg_data$trade <- relevel(factor(iti_marks_gender_cat_reg_data$trade),
                                               ref = "Electrician (NSQF)")
iti_marks_gender_cat_reg_data <- iti_marks_gender_cat_reg_data %>% 
  mutate(
    coarse_board = case_when(
      grepl("Odisha",highest_qualification_exam_board) ~ "BSE Odisha",
      grepl("Central", highest_qualification_exam_board) ~ "CBSE",
      TRUE ~ "Other"
    )
  )
iti_marks_gender_cat_reg_data$coarse_board <- relevel(factor(iti_marks_gender_cat_reg_data$coarse_board),ref = "Other")

model1 <- lm(percentage ~ gender + social_category, data = iti_marks_gender_cat_reg_data)
model2 <- lm(percentage ~ gender + social_category + trade, data=iti_marks_gender_cat_reg_data)
model3 <- lm(percentage ~ gender + social_category + coarse_board + trade, data=iti_marks_gender_cat_reg_data)
summary(model3)






```


* While the vast majority of students come from the Odisha state board, we see some other boards like CBSE as well

```{r Boards, echo=FALSE, include=TRUE}
html_table <- boards_marks %>% head(5) %>% kable(caption = "Summary statistics on marks by board") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)
  

  

```

* We can summarize average marks by social category and demographics, within different trades and different boards by running a fixed-effects regression. 

* The goal here is to *summarize* data; we do not treat the models themselves seriously.

```{r Stargazer of marks on demographics by trade, echo=FALSE, include=FALSE}
model_results <- stargazer(model1, model2, model3, omit="^trade",
          type = "html",          # Use "text" for console, "html" for HTML output in RMarkdown
          title = "Marks by demographic characteristics",
          dep.var.labels = c("Marks (%)"),
          covariate.labels = c("Female", "Green Card", "Phys. Disabled", "Scheduled Caste", "Scheduled Tribe", "BSE Odisha", "CBSE"),
          out.header = FALSE,
          align = TRUE,
          no.space = TRUE,
          star.cutoffs = c(0.05, 0.01, 0.001))


```

```{r Output stargazer of marks on demographics by trade, include=TRUE, echo=FALSE}
print_table_to_word(model_results)
```


* The estimated models have the same specifications in demographics
    * Model (2) uses within trade estimates (trade fixed-effects).
    * Model (3) = Model (2) + board fixed effects where we have classified the boards into "BSE Odisha", "CBSE", and "Other".
    * The constant in model (1) represents average marks for unreserved male students.
    * The constant in model (2) represents average marks for unreserved male students in the electrician trade.
    * The constant in model (3) represents average marks for unreserved male students in the electrician trade who graduated from some "Other" board.
* The model is missing interaction terms, which is why the coefficient estimates may not add up to within gender X category means.

### Cutoffs

Average cutoffs are low throughout ITIs 

```{r Average cutoffs by trade and demographics, echo=FALSE,include=TRUE}
html_table <- cutoffs_trade_demographics %>% head(10) %>% kable(caption = "Average cutoffs by trade and social category in 2023") %>%
    kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)


```

```{r Average cutoffs by trade and gender, echo=FALSE,include=TRUE}
html_table <- cutoffs_trade_gender %>% head(10) %>% kable(caption = "Average cutoffs by trade and social category in 2023") %>%
    kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)


```

Now we can use the fixed-effects regression approach as before to also look at how cutoffs differ between private and government ITIs.

```{r Average cutoffs by trade, gender, demographics, and type of institute, echo=FALSE,include=FALSE}

# Coarsen up social category
iti_cutoffs_2023 <- iti_cutoffs_2023 %>% mutate(
  coarse_cat = case_when(
    social_category == "SC" ~ "SC",
    social_category == "ST" ~ "ST",
    social_category == "UR" ~ "UR",
    TRUE ~ "Other"
  )
)

# Set up levels
iti_cutoffs_2023$coarse_cat <- relevel(factor(iti_cutoffs_2023$coarse_cat), ref="UR")
iti_cutoffs_2023$gender <- relevel(factor(iti_cutoffs_2023$gender), ref="Male")
iti_cutoffs_2023$type_of_institute <- relevel(factor(iti_cutoffs_2023$type_of_institute), ref="Pvt.")
iti_cutoffs_2023$trade <- relevel(factor(iti_cutoffs_2023$trade), ref= "Electrician (NSQF)")

# Estimate models
model1 <- lm(cutoff ~ type_of_institute + gender, iti_cutoffs_2023)
model2 <- lm(cutoff ~ type_of_institute + gender + coarse_cat, iti_cutoffs_2023)
model3 <- lm(cutoff ~ type_of_institute + gender + coarse_cat + trade, iti_cutoffs_2023)




```


```{r Stargazer of cutoffs on social category, gender, trade, and type, echo=FALSE, include=FALSE}
model_results <- stargazer(model1, model2, model3, omit="^trade",
          type = "html",          # Use "text" for console, "html" for HTML output in RMarkdown
          title = "Cutoffs by social category, gender, trade, and type in 2023",
          dep.var.labels = c("Marks (%)"),
          covariate.labels = c("Govt.", "Female", "Other","SC", "ST"),
          out.header = FALSE,
          align = TRUE,
          no.space = TRUE,
          star.cutoffs = c(0.05, 0.01, 0.001))
```

```{r Output Stargazer of cutoffs on social category, gender, trade, and type, include=TRUE, echo=FALSE}
print_table_to_word(model_results)

```


The uniformly low *average* cutoffs obscure the variation in cutoffs across institutes and trades. The distribution of cutoffs for general, male candidates suggests that there's substantial heterogeneity in institutional quality and trade competitiveness.

```{r Variation in cutoffs, echo=FALSE, include=TRUE}
cutoffs_ur_male <- iti_cutoffs_2023 %>% filter(coarse_cat == "UR", gender == "Male")
var_cutoffs <- sd(cutoffs_ur_male$cutoff)
var_cutoffs <- sprintf("%.1f",var_cutoffs)

ggplot(cutoffs_ur_male, aes(x = cutoff)) +
  geom_histogram(
    binwidth = 10,                 # Set bin width
    fill = "#4C72B0",             # Color for the bars
    color = "black",              # Outline color
    alpha = 0.8                   # Transparency
  ) +
  labs(
    title = "Distribution of cutoffs for general male candidates",
    x = "Cutoff (%)",
    y = "Frequency"
  ) + theme_minimal()

```


The standard deviation of the cutoff distribution is `r var_cutoffs`.



## Distance from home


```{r Load tables for distance analysis, echo=FALSE, include=FALSE}
distance_excel_path <- file.path(paths$TABLES_DIR, "distance.xlsx")
sheet_names <- getSheetNames(distance_excel_path)
distance_by_type <- read.xlsx(
  distance_excel_path,
  sheet_names[1]
)

```

```{r Summary stats on distance by type of institute, echo=FALSE,include=TRUE}
html_table <- distance_by_type %>% kable(caption="Summary stats on distance from home")  %>%
    kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)


```

Students travel slightly longer distances to private ITIs; this could be because public ITIs are even further away. It could also suggest that they are not eligible for public ITIs / trades that they are interested in.

```{r Density of distance to ITI, echo=FALSE, include=TRUE, warning=FALSE}

ggplot(iti_enrollments_2023, aes(x=distance, fill="red")) +
  geom_density(alpha=0.2) +
  theme_minimal() + scale_x_continuous(limits=c(0, 500)) +
  labs(title="Density of Distance between home and institute", x="Distance in Kilometers", y="Density") + guides(fill="none")

```

```{r Density of distance by type, echo=FALSE, include=TRUE, warning=FALSE}
ggplot(iti_enrollments_2023, aes(x=distance, fill=type_of_institute)) +
  geom_density(alpha=0.2) +
  theme_minimal() + scale_x_continuous(limits=c(0, 500)) +
  labs(title="Density of Distance between home and institute", x="Distance in Kilometers", y="Density") + labs(fill = "Type of institute")

```

We can examine the dynamics of the length of commute over time

```{r Distance over time, echo=FALSE, include=TRUE, warning=FALSE}

ggplot(iti_enrollments, aes(x = academic_year, y = distance)) +
  geom_smooth(method = "lm", se = TRUE) +  
  coord_trans(y = "log10") +
  labs(x = "Year", y = "Average distance travelled (km)") +
  theme_minimal()


```

We see a small decline in the length of commute which seems largely unrelated to a change in the number of institutes; indeed the number of institutions has remained roughly flat since 2018

```{r Number of institutes, echo=FALSE, include=TRUE, warning=FALSE}
num_institutes <- iti_enrollments %>% distinct(sams_code, academic_year, type_of_institute) %>% filter(academic_year > 2017)

ggplot(num_institutes, aes(x=academic_year, fill = type_of_institute)) +
  geom_bar()  + scale_fill_brewer("Dark2", name= "Type of institute") + labs(x = "Year", y = "Num. institutes", fill = "Type of institute") +
  theme_minimal()

```


Different social groups may be disproportionately affected by the length of the commute.

```{r Distance travelled over time by social category, echo=FALSE, include=TRUE, warning=FALSE}
distance_over_time <- iti_enrollments %>% filter(gender %in% c("Male","Female"))
ggplot(distance_over_time, aes(x = academic_year, y = distance, color = social_category, group = social_category)) +
  geom_smooth(method = "lm", se = TRUE) +  
  coord_trans(y = "log10") +
  labs(x = "Year", y = "Average distance travelled (km)", color = "Social category") +
  theme_minimal()

```


```{r Distance travelled over time by gender, echo=FALSE, include=TRUE, warning=FALSE}
ggplot(distance_over_time, aes(x = academic_year, y = distance, color = gender, group = gender)) +
  geom_smooth(method = "lm", se = TRUE) +  
  coord_trans(y = "log10") +
  labs(x = "Year", y = "Average distance travelled (km)", color = "Gender") +
  theme_minimal()
```


## Over/undersubscription

```{r Load vacancies tables, echo=FALSE, include=FALSE}
vacancy_path <- file.path(paths$TABLES_DIR, "vacancies.xlsx") 
sheet_names <- getSheetNames(vacancy_path)

vacancy_by_trade <- read.xlsx(
  vacancy_path,
  sheet_names[1]
)

vacancy_ratio_by_trade <- read.xlsx(
  vacancy_path,
  sheet_names[2]
)

top_10_institutes_by_vacancy <- read.xlsx(
  vacancy_path,
  sheet_names[3]
)

bottom_10_institutes_by_vacancy <- read.xlsx(
  vacancy_path,
  sheet_names[4]
)

top_10_institutes_by_vacancy_ratio <- read.xlsx(
  vacancy_path,
  sheet_names[5]
)

bottom_10_institutes_by_vacancy_ratio <- read.xlsx(
  vacancy_path,
  sheet_names[6]
)

top_10_districts_by_vacancy <- read.xlsx(
  vacancy_path,
  sheet_names[7]
)

bottom_10_districts_by_vacancy <- read.xlsx(
  vacancy_path,
  sheet_names[8]
)

top_10_districts_by_vacancy_ratio <- read.xlsx(
  vacancy_path,
  sheet_names[9]
)

bottom_10_districts_by_vacancy_ratio <- read.xlsx(
  vacancy_path,
  sheet_names[10]
)

top_10_trades_by_vacancy <- read.xlsx(
  vacancy_path,
  sheet_names[11]
)

bottom_10_trades_by_vacancy <- read.xlsx(
  vacancy_path,
  sheet_names[12]
)

top_10_trades_by_vacancy_ratio <- read.xlsx(
  vacancy_path,
  sheet_names[13]
)

bottom_10_trades_by_vacancy_ratio <- read.xlsx(
  vacancy_path,
  sheet_names[14]
)



```

```{r Vacancy by trade, echo=FALSE, include=TRUE}

html_table <- vacancy_by_trade %>% kable(caption = "Summary stats on vacancies by trade in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)
  


```


```{r Vacancy ratio by trade, echo=FALSE, include=TRUE}

html_table <- vacancy_ratio_by_trade %>% kable(caption = "Summary stats on vacancy ratios by trade in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)


```

```{r Top 10 institutes by vacancy, echo=FALSE, include=TRUE}
html_table <- top_10_institutes_by_vacancy %>% kable(caption = "Top 10 institutes by vacancy in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)

```

```{r Bottom 10 institutes by vacancy, echo=FALSE, include=TRUE}

html_table <- bottom_10_institutes_by_vacancy %>% kable(caption = "Bottom 10 institutes by vacancy in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)

```


```{r Top 10 institutes by vacancy ratio, echo=FALSE, include=TRUE}
html_table <- top_10_institutes_by_vacancy_ratio %>% kable(caption = "Top 10 institutes by vacancy ratio in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)

```


```{r Bottom 10 institutes by vacancy ratio, echo=FALSE, include=TRUE}

html_table <- bottom_10_institutes_by_vacancy_ratio %>% kable(caption = "Bottom 10 institutes by vacancy ratio in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)

```


```{r Top 10 districts by vacancy, echo=FALSE, include=TRUE}
html_table <- top_10_districts_by_vacancy %>% kable(caption = "Top 10 districts by vacancy in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)

```

```{r Bottom 10 districts by vacancy, echo=FALSE, include=TRUE}

html_table <- bottom_10_districts_by_vacancy %>% kable(caption = "Bottom 10 districts by vacancy in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)

```


```{r Top 10 districts by vacancy ratio, echo=FALSE, include=TRUE}
html_table <- top_10_districts_by_vacancy_ratio %>% kable(caption = "Top 10 districts by vacancy ratio in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)

```


```{r Bottom 10 districts by vacancy ratio, echo=FALSE, include=TRUE}

html_table <- bottom_10_districts_by_vacancy_ratio %>% kable(caption = "Bottom 10 districts by vacancy ratio in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)

```



```{r Top 10 trades by vacancy, echo=FALSE, include=TRUE}
html_table <- top_10_trades_by_vacancy %>% kable(caption = "Top 10 trades by vacancy in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)

```

```{r Bottom 10 trades by vacancy, echo=FALSE, include=TRUE}

html_table <- bottom_10_trades_by_vacancy %>% kable(caption = "Bottom 10 trades by vacancy in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)

```


```{r Top 10 trades by vacancy ratio, echo=FALSE, include=TRUE}
html_table <- top_10_trades_by_vacancy_ratio %>% kable(caption = "Top 10 trades by vacancy ratio in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)

```


```{r Bottom 10 trades by vacancy ratio, echo=FALSE, include=TRUE}

html_table <- bottom_10_trades_by_vacancy_ratio %>% kable(caption = "Bottom 10 trades by vacancy ratio in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)

```


## Geography

```{r Load geography tables, echo=TRUE, include=FALSE}

geography <- file.path(paths$TABLES_DIR,"geography.xlsx")
sheet_names <- getSheetNames(geography)
female_share_by_district <- read.xlsx(
  geography,
  sheet_names[1]
) %>% rename(`Female share` = Female.share)

```


```{r Female share by district, echo=TRUE, include=TRUE}
html_table <- female_share_by_district %>% kable(caption = "Female share by district") %>%
    kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

print_table_to_word(html_table)


```








