---
title: "SAMS Data Analysis"
author: "Data, Policy, and Innovation Centre"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    theme: journal
    mathjax: default
    toc: true
    toc_depth: 3
    fig_caption: true
    keep_md: false           # Choose whether to keep the intermediate .md file
    df_print: paged         # Prints data frames as paginated tables
---

```{r Load configuration, include=FALSE, eval=FALSE, echo=FALSE}
library(here)
rm(list = ls())

here::i_am("README.md")
here()
source(file.path(here(),"scripts","config.R"))


```

```{r Libraries, include=FALSE, echo=FALSE}
library(tidyverse)
library(arrow)
library(ggplot2)
library(openxlsx)
library(kableExtra)
library(knitr)
library(stargazer)
library(RColorBrewer)
```

```{r Load processed data, include=FALSE, echo=FALSE}
# Processed data
iti_marks_and_cutoffs <- read_parquet(get_path("iti_marks_and_cutoffs"))
iti_vacancies <- read_parquet(get_path("iti_vacancies"))
iti_marks_and_cutoffs_2023 <- iti_marks_and_cutoffs %>% filter(academic_year == 2023)
```


```{r Load interim data, include=FALSE, echo=FALSE}
# Interim data
iti_enrollments <- read_parquet(get_path("iti_enrollments"))
iti_enrollments_2023 <- iti_enrollments %>% filter(academic_year == 2023)
iti_cutoffs <- read_parquet(get_path("iti_institutes_cutoffs"))

```

```{r Add type of institutes data to cutoffs, include=FALSE, echo=FALSE}
iti_cutoffs <- iti_enrollments %>% distinct(sams_code,type_of_institute) %>% 
  right_join(iti_cutoffs, by="sams_code")
iti_cutoffs_2023 <- iti_cutoffs %>% filter(academic_year == 2023)


```

# Executive summary

# An overview of SAMS

## Enrollment

## Institutes

## Trades

# The vocational training pipeline

# Student demographics

# Student qualifications and marks

# The geography of vocational training








<!-- # Goals for the report
The long-term goal of this analysis is to uncover motivations and patterns behind students’ enrollment in vocational training program and uncover factors influencing choice based on gender and social demographics. Of principal interest to us is

* What types of students decide to enroll in vocational training programs?
    * This ideally requires data on enrollment in higher education / 12th grade programs as well

* Are there "overqualified" students enrolling in vocational programs?

* Does distance from training institutes act as a binding constraint on students?
    * Are some students unable to enroll in the trades / institutions they would like because of a long commute?
    * Hostel facilities are likely to play a role in this.

* Are there certain trades that are "oversubscribed", in that the government has not provided enough seats in those trades despite demand?
    * Conversely, are some other trades "undersubscribed", and perhaps resources are better allocated in other trades?

The following exhibits serve as a scoping exercise for the report due at the end of the month. -->

# Raw exhibits

## Trades and qualifications

```{r Load output tables, include=FALSE, echo=FALSE}
quals_excel_path <- file.path(paths$TABLES_DIR,"trades_by_qualification.xlsx")
sheet_names <- getSheetNames(quals_excel_path)
trades_by_quals <- read.xlsx(
  quals_excel_path,
  sheet_names[1]
)
trades_by_quals_pct <- read.xlsx(
  quals_excel_path,
  sheet_names[2]
) %>% rename(
  `Graduate and above` = Graduate.and.above,
  Trade = reported_branch_or_trade
)
trades_by_gender_pct <- read.xlsx(
  quals_excel_path,
  sheet_names[3]
) %>% rename(
  Trade = reported_branch_or_trade
)

qual_by_gender_pct <- read.xlsx(
   quals_excel_path,
   sheet_names[4]
) %>% rename(
  `Graduate and above` = Graduate.and.above
)

trades_by_soc_pct <- read.xlsx(
  quals_excel_path,
  sheet_names[5]
) %>% rename(
  Trade = reported_branch_or_trade
)

qual_by_soc_pct <- read.xlsx(
  quals_excel_path,
  sheet_names[6]
) %>% 
  rename(
    `Graduate and above` = Graduate.and.above
  )

```


* We are likely to find some trades where students are "overqualified". 

```{r Highest qualifications raw, include=FALSE, echo=FALSE}

 trades_by_quals %>% kable( caption = "Trades by highest qualifications") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))
```

```{r Highest qualifications shares, include=TRUE, echo=FALSE}
trades_by_quals_pct %>% head(10) %>% kable( caption = "Top 10 trades with highest share of college graduates") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

```
* Some trades show a particularly high share of graduates
    * These appear to be female dominated trades
    
```{r Trades by gender, include=TRUE, echo=FALSE}
top_10_trades_by_quals_grad <- trades_by_quals_pct %>% head(10)
trades_by_gender_pct %>% right_join(top_10_trades_by_quals_grad,by="Trade") %>% select(Trade,Male,Female, Total.x) %>% rename(Total = Total.x) %>% kable( caption = "Top 10 trades by highest share of college graduates, sorted by gender (%)") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))


```

```{r Qualification by gender, echo=FALSE, include=TRUE}
qual_by_gender_pct %>%  kable( caption = "Highest qualification by gender (%)") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

```


```{r Trades by social category, echo=FALSE, include=TRUE}
trades_by_soc_pct %>% head(20) %>%  kable( caption = "Trades by social category in 2023 (%)") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

```
 * Need to check if heavy sorting into trades is due to the geography of trade offerings.
 
```{r Qualifications by social category, echo=FALSE, include=TRUE}
qual_by_soc_pct %>% kable( caption = "Qualifications by social category in 2023 (%)") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))


```
 

## Student marks and institute cutoffs

### Marks

```{r Load tables for marks analysis, echo=FALSE,include=FALSE}
marks_excel_path <- file.path(paths$TABLES_DIR, "marks_cutoffs.xlsx")
sheet_names <- getSheetNames(marks_excel_path)
marks_by_trade <- read.xlsx(
  marks_excel_path,
  sheet_names[1]
  
)
marks_by_demographics <- read.xlsx(
  marks_excel_path,
  sheet_names[2]
)
boards_marks <- read.xlsx(
  marks_excel_path,
  sheet_names[3]
)

cutoffs_trade_demographics <- read.xlsx(
  marks_excel_path,
  sheet_names[4]
)

cutoffs_trade_gender <- read.xlsx(
  marks_excel_path,
  sheet_names[5]
)

```

```{r Distribution of 10th marks, echo=FALSE, include=TRUE, warning=FALSE}

ggplot(iti_marks_and_cutoffs_2023, aes(x = percentage)) +
  geom_histogram(
    binwidth = 1,                 # Set bin width
    fill = "#4C72B0",             # Color for the bars
    color = "black",              # Outline color
    alpha = 0.8                   # Transparency
  ) +
  labs(
    title = "Overall distribution of marks in 2023",
    x = "Marks (%)",
    y = "Frequency"
  ) + 
  geom_vline(
    xintercept = 33,              # Set the x position of the line at 33 marks
    color = "red",                # Color of the line
    linetype = "dashed",          # Line type (dashed)
    size = 1                      # Thickness of the line
  ) + 
  annotate(
    "text", 
    x = 33, y = 4500,                # Position the text below the line
    label = "Passing Percentage: 33%", 
    color = "red",                # Color of the text
    size = 3,                     # Size of the text
    hjust = -0.05,
    vjust = -1
  
  ) + 
  theme_classic() 
  

```

```{r Average marks by trade, echo=FALSE, include=TRUE}
marks_by_trade %>% head(20) %>% kable( caption = "Summary statistics on marks by trade") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

```

```{r Average marks by demographics, echo=FALSE,include=TRUE}
marks_by_demographics %>% kable(caption = "Average marks by demographics") %>% 
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))
  

```



```{r Average marks by gender and category within trade, echo=FALSE, include=FALSE}

iti_marks_gender_cat_reg_data <- iti_marks_and_cutoffs_2023
iti_marks_gender_cat_reg_data$gender <- relevel(factor(iti_marks_gender_cat_reg_data$gender),ref="Male")

iti_marks_gender_cat_reg_data$social_category <- relevel(factor(iti_marks_gender_cat_reg_data$social_category),ref="UR")

iti_marks_gender_cat_reg_data$trade <- relevel(factor(iti_marks_gender_cat_reg_data$trade),
                                               ref = "Electrician (NSQF)")
iti_marks_gender_cat_reg_data <- iti_marks_gender_cat_reg_data %>% 
  mutate(
    coarse_board = case_when(
      grepl("Odisha",highest_qualification_exam_board) ~ "BSE Odisha",
      grepl("Central", highest_qualification_exam_board) ~ "CBSE",
      TRUE ~ "Other"
    )
  )
iti_marks_gender_cat_reg_data$coarse_board <- relevel(factor(iti_marks_gender_cat_reg_data$coarse_board),ref = "Other")

model1 <- lm(percentage ~ gender + social_category, data = iti_marks_gender_cat_reg_data)
model2 <- lm(percentage ~ gender + social_category + trade, data=iti_marks_gender_cat_reg_data)
model3 <- lm(percentage ~ gender + social_category + coarse_board + trade, data=iti_marks_gender_cat_reg_data)
summary(model3)






```


* While the vast majority of students come from the Odisha state board, we see some other boards like CBSE as well

```{r Boards, echo=FALSE, include=TRUE}
boards_marks %>% head(5) %>% kable(caption = "Summary statistics on marks by board") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))
  

  

```

* We can summarize average marks by social category and demographics, within different trades and different boards by running a fixed-effects regression. 

* The goal here is to *summarize* data; we do not treat the models themselves seriously.

```{r Stargazer of marks on demographics by trade, echo=FALSE, results='asis'}
model_results <- stargazer(model1, model2, model3, omit="^trade",
          type = "html",          # Use "text" for console, "html" for HTML output in RMarkdown
          title = "Marks by demographic characteristics",
          dep.var.labels = c("Marks (%)"),
          covariate.labels = c("Female", "Green Card", "Phys. Disabled", "Scheduled Caste", "Scheduled Tribe", "BSE Odisha", "CBSE"),
          out.header = FALSE,
          align = TRUE,
          no.space = TRUE,
          star.cutoffs = c(0.05, 0.01, 0.001))
```

* The estimated models have the same specifications in demographics
    * Model (2) uses within trade estimates (trade fixed-effects).
    * Model (3) = Model (2) + board fixed effects where we have classified the boards into "BSE Odisha", "CBSE", and "Other".
    * The constant in model (1) represents average marks for unreserved male students.
    * The constant in model (2) represents average marks for unreserved male students in the electrician trade.
    * The constant in model (3) represents average marks for unreserved male students in the electrician trade who graduated from some "Other" board.
* The model is missing interaction terms, which is why the coefficient estimates may not add up to within gender X category means.

### Cutoffs

Average cutoffs are low throughout ITIs 

```{r Average cutoffs by trade and demographics, echo=FALSE,include=TRUE}
cutoffs_trade_demographics %>% head(10) %>% kable(caption = "Average cutoffs by trade and social category in 2023") %>%
    kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))


```

```{r Average cutoffs by trade and gender, echo=FALSE,include=TRUE}
cutoffs_trade_gender %>% head(10) %>% kable(caption = "Average cutoffs by trade and social category in 2023") %>%
    kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))


```

Now we can use the fixed-effects regression approach as before to also look at how cutoffs differ between private and government ITIs.

```{r Average cutoffs by trade, gender, demographics, and type of institute, echo=FALSE,include=FALSE}

# Coarsen up social category
iti_cutoffs_2023 <- iti_cutoffs_2023 %>% mutate(
  coarse_cat = case_when(
    social_category == "SC" ~ "SC",
    social_category == "ST" ~ "ST",
    social_category == "UR" ~ "UR",
    TRUE ~ "Other"
  )
)

# Set up levels
iti_cutoffs_2023$coarse_cat <- relevel(factor(iti_cutoffs_2023$coarse_cat), ref="UR")
iti_cutoffs_2023$gender <- relevel(factor(iti_cutoffs_2023$gender), ref="Male")
iti_cutoffs_2023$type_of_institute <- relevel(factor(iti_cutoffs_2023$type_of_institute), ref="Pvt.")
iti_cutoffs_2023$trade <- relevel(factor(iti_cutoffs_2023$trade), ref= "Electrician (NSQF)")

# Estimate models
model1 <- lm(cutoff ~ type_of_institute + gender, iti_cutoffs_2023)
model2 <- lm(cutoff ~ type_of_institute + gender + coarse_cat, iti_cutoffs_2023)
model3 <- lm(cutoff ~ type_of_institute + gender + coarse_cat + trade, iti_cutoffs_2023)




```


```{r Stargazer of cutoffs on social category, gender, trade, and type, echo=FALSE, results='asis'}
model_results <- stargazer(model1, model2, model3, omit="^trade",
          type = "html",          # Use "text" for console, "html" for HTML output in RMarkdown
          title = "Cutoffs by social category, gender, trade, and type in 2023",
          dep.var.labels = c("Marks (%)"),
          covariate.labels = c("Govt.", "Female", "Other","SC", "ST"),
          out.header = FALSE,
          align = TRUE,
          no.space = TRUE,
          star.cutoffs = c(0.05, 0.01, 0.001))
```




The uniformly low *average* cutoffs obscure the variation in cutoffs across institutes and trades. The distribution of cutoffs for general, male candidates suggests that there's substantial heterogeneity in institutional quality and trade competitiveness.

```{r Variation in cutoffs, echo=FALSE, include=TRUE}
cutoffs_ur_male <- iti_cutoffs_2023 %>% filter(coarse_cat == "UR", gender == "Male")
var_cutoffs <- sd(cutoffs_ur_male$cutoff)
var_cutoffs <- sprintf("%.1f",var_cutoffs)

ggplot(cutoffs_ur_male, aes(x = cutoff)) +
  geom_histogram(
    binwidth = 10,                 # Set bin width
    fill = "#4C72B0",             # Color for the bars
    color = "black",              # Outline color
    alpha = 0.8                   # Transparency
  ) +
  labs(
    title = "Distribution of cutoffs for general male candidates",
    x = "Cutoff (%)",
    y = "Frequency"
  ) + theme_minimal()

```


The standard deviation of the cutoff distribution is `r var_cutoffs`.



## Distance from home


```{r Load tables for distance analysis, echo=FALSE, include=FALSE}
distance_excel_path <- file.path(paths$TABLES_DIR, "distance.xlsx")
sheet_names <- getSheetNames(distance_excel_path)
distance_by_type <- read.xlsx(
  distance_excel_path,
  sheet_names[1]
)

```

```{r Summary stats on distance by type of institute, echo=FALSE,include=TRUE}
distance_by_type %>% kable(caption="Summary stats on distance from home")  %>%
    kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))


```

Students travel slightly longer distances to private ITIs; this could be because public ITIs are even further away. It could also suggest that they are not eligible for public ITIs / trades that they are interested in.

```{r Density of distance to ITI, echo=FALSE, include=TRUE, warning=FALSE}

ggplot(iti_enrollments_2023, aes(x=distance, fill="red")) +
  geom_density(alpha=0.2) +
  theme_minimal() + scale_x_continuous(limits=c(0, 500)) +
  labs(title="Density of Distance between home and institute", x="Distance in Kilometers", y="Density") + guides(fill="none")

```

```{r Density of distance by type, echo=FALSE, include=TRUE, warning=FALSE}
ggplot(iti_enrollments_2023, aes(x=distance, fill=type_of_institute)) +
  geom_density(alpha=0.2) +
  theme_minimal() + scale_x_continuous(limits=c(0, 500)) +
  labs(title="Density of Distance between home and institute", x="Distance in Kilometers", y="Density") + labs(fill = "Type of institute")

```

We can examine the dynamics of the length of commute over time

```{r Distance over time, echo=FALSE, include=TRUE, warning=FALSE}

ggplot(iti_enrollments, aes(x = academic_year, y = distance)) +
  geom_smooth(method = "lm", se = TRUE) +  
  coord_trans(y = "log10") +
  labs(x = "Year", y = "Average distance travelled (km)") +
  theme_minimal()


```

We see a small decline in the length of commute which seems largely unrelated to a change in the number of institutes; indeed the number of institutions has remained roughly flat since 2018

```{r Number of institutes, echo=FALSE, include=TRUE, warning=FALSE}
num_institutes <- iti_enrollments %>% distinct(sams_code, academic_year, type_of_institute) %>% filter(academic_year > 2017)

ggplot(num_institutes, aes(x=academic_year, fill = type_of_institute)) +
  geom_bar()  + scale_fill_brewer("Dark2", name= "Type of institute") + labs(x = "Year", y = "Num. institutes", fill = "Type of institute") +
  theme_minimal()

```


Different social groups may be disproportionately affected by the length of the commute.

```{r Distance travelled over time by social category, echo=FALSE, include=TRUE, warning=FALSE}
distance_over_time <- iti_enrollments %>% filter(gender %in% c("Male","Female"))
ggplot(distance_over_time, aes(x = academic_year, y = distance, color = social_category, group = social_category)) +
  geom_smooth(method = "lm", se = TRUE) +  
  coord_trans(y = "log10") +
  labs(x = "Year", y = "Average distance travelled (km)", color = "Social category") +
  theme_minimal()

```


```{r Distance travelled over time by gender, echo=FALSE, include=TRUE, warning=FALSE}
ggplot(distance_over_time, aes(x = academic_year, y = distance, color = gender, group = gender)) +
  geom_smooth(method = "lm", se = TRUE) +  
  coord_trans(y = "log10") +
  labs(x = "Year", y = "Average distance travelled (km)", color = "Gender") +
  theme_minimal()
```


## Over/undersubscription

```{r Load vacancies tables, echo=FALSE, include=FALSE}
vacancy_path <- file.path(paths$TABLES_DIR, "vacancies.xlsx") 
sheet_names <- getSheetNames(vacancy_path)

vacancy_by_trade <- read.xlsx(
  vacancy_path,
  sheet_names[1]
)

vacancy_ratio_by_trade <- read.xlsx(
  vacancy_path,
  sheet_names[2]
)

top_10_institutes_by_vacancy <- read.xlsx(
  vacancy_path,
  sheet_names[3]
)

bottom_10_institutes_by_vacancy <- read.xlsx(
  vacancy_path,
  sheet_names[4]
)

top_10_institutes_by_vacancy_ratio <- read.xlsx(
  vacancy_path,
  sheet_names[5]
)

bottom_10_institutes_by_vacancy_ratio <- read.xlsx(
  vacancy_path,
  sheet_names[6]
)

top_10_districts_by_vacancy <- read.xlsx(
  vacancy_path,
  sheet_names[7]
)

bottom_10_districts_by_vacancy <- read.xlsx(
  vacancy_path,
  sheet_names[8]
)

top_10_districts_by_vacancy_ratio <- read.xlsx(
  vacancy_path,
  sheet_names[9]
)

bottom_10_districts_by_vacancy_ratio <- read.xlsx(
  vacancy_path,
  sheet_names[10]
)

top_10_trades_by_vacancy <- read.xlsx(
  vacancy_path,
  sheet_names[11]
)

bottom_10_trades_by_vacancy <- read.xlsx(
  vacancy_path,
  sheet_names[12]
)

top_10_trades_by_vacancy_ratio <- read.xlsx(
  vacancy_path,
  sheet_names[13]
)

bottom_10_trades_by_vacancy_ratio <- read.xlsx(
  vacancy_path,
  sheet_names[14]
)



```

```{r Vacancy by trade, echo=FALSE, include=TRUE}

vacancy_by_trade %>% kable(caption = "Summary stats on vacancies by trade in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))
  


```


```{r Vacancy ratio by trade, echo=FALSE, include=TRUE}

vacancy_ratio_by_trade %>% kable(caption = "Summary stats on vacancy ratios by trade in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))


```

```{r Top 10 institutes by vacancy, echo=FALSE, include=TRUE}
top_10_institutes_by_vacancy %>% kable(caption = "Top 10 institutes by vacancy in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

```

```{r Bottom 10 institutes by vacancy, echo=FALSE, include=TRUE}

bottom_10_institutes_by_vacancy %>% kable(caption = "Bottom 10 institutes by vacancy in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

```


```{r Top 10 institutes by vacancy ratio, echo=FALSE, include=TRUE}
top_10_institutes_by_vacancy_ratio %>% kable(caption = "Top 10 institutes by vacancy ratio in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

```


```{r Bottom 10 institutes by vacancy ratio, echo=FALSE, include=TRUE}

bottom_10_institutes_by_vacancy_ratio %>% kable(caption = "Bottom 10 institutes by vacancy ratio in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

```


```{r Top 10 districts by vacancy, echo=FALSE, include=TRUE}
top_10_districts_by_vacancy %>% kable(caption = "Top 10 districts by vacancy in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

```

```{r Bottom 10 districts by vacancy, echo=FALSE, include=TRUE}

bottom_10_districts_by_vacancy %>% kable(caption = "Bottom 10 districts by vacancy in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

```


```{r Top 10 districts by vacancy ratio, echo=FALSE, include=TRUE}
top_10_districts_by_vacancy_ratio %>% kable(caption = "Top 10 districts by vacancy ratio in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

```


```{r Bottom 10 districts by vacancy ratio, echo=FALSE, include=TRUE}

bottom_10_districts_by_vacancy_ratio %>% kable(caption = "Bottom 10 districts by vacancy ratio in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

```



```{r Top 10 trades by vacancy, echo=FALSE, include=TRUE}
top_10_trades_by_vacancy %>% kable(caption = "Top 10 trades by vacancy in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

```

```{r Bottom 10 trades by vacancy, echo=FALSE, include=TRUE}

bottom_10_trades_by_vacancy %>% kable(caption = "Bottom 10 trades by vacancy in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

```


```{r Top 10 trades by vacancy ratio, echo=FALSE, include=TRUE}
top_10_trades_by_vacancy_ratio %>% kable(caption = "Top 10 trades by vacancy ratio in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

```


```{r Bottom 10 trades by vacancy ratio, echo=FALSE, include=TRUE}

bottom_10_trades_by_vacancy_ratio %>% kable(caption = "Bottom 10 trades by vacancy ratio in 2023") %>%
  kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))

```


## Geography

```{r Load geography tables, echo=TRUE, include=FALSE}

geography <- file.path(paths$TABLES_DIR,"geography.xlsx")
sheet_names <- getSheetNames(geography)
female_share_by_district <- read.xlsx(
  geography,
  sheet_names[1]
) %>% rename(`Female share` = Female.share)

```


```{r Female share by district, echo=TRUE, include=TRUE}
female_share_by_district %>% kable(caption = "Female share by district") %>%
    kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))


```


## Price

TODO -- Need to collect more price information







